### 2012-11-21 Publications/2013-IPMI/Latex/02-Abstract.tex
Added base for IPMI2013 #25 I upload here the introduction. I take the last paragraph almost literally from the original Dominique's document.

\begin{abstract}
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words. It will be set in 9-point
font size and be inset 1.0 cm from the right and left margins.
There will be two blank lines before and after the Abstract.
%\dots
%\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
\end{abstract}