### 2012-11-21 Publications/2013-IPMI/Latex/05-Results.tex

\section{Results and discussion}
\label{sec:results}
%