### 2012-12-15 Publications/2013-IPMI/Latex/00-Main.tex
fixing sqrt in methods

\documentclass{llncs}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{acronym}
\usepackage[numbers]{natbib}
\bibliographystyle{splncs03}

\usepackage{nameref}
\usepackage{hyperref}
%\usepackage[hyphens]{url}
%\usepackage[hyphenbreaks]{breakurl}
\usepackage[USenglish]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pslatex}
%
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}

\usepackage{graphicx}
%
%\tableofcontents
%
\mainmatter              % start of the contributions
%
\graphicspath{{../Figures/}}

\begin{document}

\title{Simultaneous segmentation and distortion correction on diffusion weighted MR using shape priors}
%%
\titlerunning{Simultaneous segmentation and distortion correction on DWI}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Oscar Esteban\inst{1,2} \and
        Alessandro Daducci \inst{2} \and
		Meritxell Bach-Cuadra \inst{2,3} \and
        Jean-Philippe Thiran \inst{2,3} \and
        Andr\'es Santos \inst{1} \and
        Dominique Zosso \inst{2,4}}
%
\authorrunning{Oscar Esteban et al.} % abbreviated author list (for running head)

\institute{Biomedical Image Technologies (BIT),
ETSI Telecomunicaci\'on - Universidad Polit\'ecnica de Madrid
and CIBER-BBN, \\
Av. Complutense 30, E-28040 Madrid, Spain\\
\email{phd@oscaresteban.es},\\
% WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
\and
Signal Processing Laboratory (LTS5),
\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
EPFL STI IEL LTS5, Station 11, CH-1015 Lausanne, Switzerland
\and
Dept. of Radiology, University Hospital Center (CHUV) and University of Lausanne (UNIL) \\
Rue du Bugnon 46, CH-1011 Lausanne, Switzerland
\and
Department of Mathematics, 
University of California, Los Angeles (UCLA) \\
520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, USA
}

\maketitle              % typeset the title of the contribution

\input{01-Acronyms}

\begin{abstract}
\input{02-Abstract}
\keywords{ magnetic resonance, diffusion weighted imaging, distortion correction, segmentation, registration, shape priors, connectomics, echo planar imaging
           }
\end{abstract}

\input{03-Introduction}

\input{04-Methods}

\input{045-Experiments}

\input{05-Results}

\input{06-Conclusion}

%\input{07-Acknowledgments}

\bibliography{99-References}

\end{document}