### 2012-12-22 Publications/2013-IPMI/Latex/00-Main.tex
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

\documentclass{llncs}
\usepackage{makeidx}  % allows for indexgeneration
%\usepackage{acronym}
%\usepackage{acro}
\usepackage{glossaries}
\usepackage[numbers]{natbib}
\bibliographystyle{splncs03}

\usepackage{nameref}
\usepackage{hyperref}
%\usepackage[hyphens]{url}
%\usepackage[hyphenbreaks]{breakurl}
\usepackage[USenglish]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pslatex}
%\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{xcolor}

%\renewcommand{\arraystretch}{0.8}
%
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\usepackage{graphicx}

%
%\tableofcontents
%
\mainmatter              % start of the contributions
%
\graphicspath{{../Figures/}}

\input{01-Acronyms}

\begin{document}

\title{Simultaneous segmentation and distortion correction on diffusion weighted MR using shape priors}
%%
\titlerunning{Simultaneous segmentation and distortion correction on DWI}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
%\author{Oscar Esteban\inst{1,2} \and
%        Alessandro Daducci \inst{2} \and
%		Meritxell Bach-Cuadra \inst{2,3} \and
%        Jean-Philippe Thiran \inst{2,3} \and
%        Andr\'es Santos \inst{1} \and
%        Dominique Zosso \inst{2,4}}
%
%\authorrunning{Oscar Esteban et al.} % abbreviated author list (for running head)

%\institute{Biomedical Image Technologies (BIT),
%ETSI Telecomunicaci\'on - Universidad Polit\'ecnica de Madrid
%and CIBER-BBN, \\
%Av. Complutense 30, E-28040 Madrid, Spain\\
%\email{phd@oscaresteban.es},\\
%\and
%Signal Processing Laboratory (LTS5),
%\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
%EPFL STI IEL LTS5, Station 11, CH-1015 Lausanne, Switzerland
%\and
%Dept. of Radiology, University Hospital Center (CHUV) and University of Lausanne (UNIL) \\
%Rue du Bugnon 46, CH-1011 Lausanne, Switzerland
%\and
%Department of Mathematics, 
%University of California, Los Angeles (UCLA) \\
%520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, USA
%}

\maketitle              % typeset the title of the contribution

\begin{abstract}
\input{02-Abstract}
\keywords{ magnetic resonance, diffusion weighted imaging, distortion correction, segmentation, registration, shape priors, connectomics, echo planar imaging
           }
\end{abstract}

\input{03-Introduction}

\input{04-Methods}

\input{045-Experiments}

\input{05-Results}

\input{06-Conclusion}

%\input{07-Acknowledgments}

%\bibliography{99-References}
\bibliography{98-References-short}

\end{document}