### 2013-05-17 Publications/2013-TMI/08-appendix-01.tex
Updated files. Copied everything from IPMI
Also updated references to add one to the new PlosOne
article.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices

\section{Proof of duality between the probabilistic and the presented frameworks}
Here the demonstration that we are preparing.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.