### 2013-05-20 Publications/2013-TMI/08-appendix-01.tex

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices

\section{Proof of duality between the probabilistic and the presented frameworks}
Here the demonstration that we are preparing.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

In the probabilistic approach to image segmentation, the smoothness constraints
can be introduced into the \gls{map} criterium as a \gls{mrf}. Accordingly to the
Hammersley-Clifford theorem (\textbf{FIXME}: citations needed), an \gls{mrf} can
be equivalently characterized by a Gibb's distribution:
\begin{equation}
P(\mathbf{K})=\underset{i}{\prod} Z(U)^{-1}\,e^{-U(\mathbf{x}_i \mid \beta,K)} = \mathit{const.}\,e^{- \sum\limits_i \left( U(\mathbf{x}_i \mid \beta,K) \right) } \sim e^{-\nu_B \left| K \right| }.
\end{equation}
This way, we include the assumption that the total length of the edge set $K$ is small
and we draw the equivalence with the \gls{mrf} modeling, for $U(\mathbf{x}_i \mid \beta,K)$
being the Potts model.

In the probabilistic approach, the underlying concept to segmenting an image is the
Bayes' rule. For a discrete image of $N$ pixels indexed by $i$, this rule reads as
follows:
\begin{equation}
P(K|I)=P(I \mid K)\, P(K)=\underset{i}{\prod} p (\mathbf{x}_i | K ) \, p_i(K).
\end{equation}
The normalizer probability $P(I)$ has been ommited for simplicity.

In order to express the likelihood as an absolute energy in our variation framework,
the \textit{log}-likelihood is computed:
\begin{equation}
E(K)= -\log( P(K|I) ) = \sum\limits_k \int_{\Omega_k} -\log p_k(I(\mathbf{x}),\mathbf{x}) \,d\mathbf{x}+\nu_B \left|K\right|,
\end{equation}
where $\mathbf{s} = I(\mathbf{x})$, $\mathbf{s} \in \mathbb{R}^C$ (a vector of $C$ scalar features).
Additionally, the discrete grid of $N$ pixels has been converted to a continuous space by assuming
$d\mathbf{x}$ as infinitesimal bin size.

To explicitly define the likelihood, we firstly define the squared \textit{Mahalanobis distance} that
is the exponential of a multivariate normal distribution:
\begin{equation}
\Delta^2_k (\mathbf{s}) = (\mathbf{s} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{s} - \boldsymbol{\mu}_k).
\end{equation}

Therefore, the likelihood is defined as follows:
\begin{equation}
p_k(I(\mathbf{x}),\mathbf{x}) = p_k(\mathbf{s},\mathbf{x})= p_k(\mathbf{s}) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{s}) \right)}}.
\end{equation}

Introducing this definition on (14), we have:
\begin{equation}
E(K)= \sum\limits_k \int_{\Omega_k} -\log{\left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{s}) \right)} \right] } \,d\mathbf{x}+\nu_B \left|K\right|.
\end{equation}

\begin{equation}
E(K) = \sum\limits_k \int_{\Omega_k} \left( \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{s}) \right) \,d\mathbf{x}+\nu_B \left|K\right|
\end{equation}

\begin{equation}
E(K) = \sum\limits_k \left( \frac{ V_k }{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)}+ \frac{1}{2} \int_{\Omega_k} \Delta^2_k (\mathbf{s}) \,d\mathbf{x}+\nu_B \left|K\right| \right)
\end{equation}