### 2013-06-04 Publications/2013-FINF-ITK/01-abstract.tex

In whole-brain connectivity analysis of diffusion MRI (dMRI) data,
an accurate delineation of the white-matter and grey-matter surfaces is required.
While high-standard segmentation is readily available for anatomical MRI, such 
as T1-weighted, dMRI typically have lower resolution and severe 
geometrical distortions. We propose a dMRI segmentation-registration framework 
that exploits the detailed anatomy extracted from anatomical MRI as shape-prior.
We use an ``active contours without edges''-like model to look for a deformation
field that optimally maps the shape prior on the multivariate features 
in diffusion space. This joint approach reflects the intrinsic coupling of 
segmentation and distortion correction. Complementary, a precise and consistent
cortical parcellation on dMRI is straightforward by projection from T1 space. Thus,
we expect to improve the reliability and robustness of the resulting connectivity 
networks and their comparability within and across subjects. First results 
on synthetic datasets and simulated dMRI confirm the effectiveness of our approach.