### 2013-09-19 Publications/2013-ISBI14/01-abstract.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

\begin{abstract}

Connectivity analysis on \gls*{dmri} data of the whole-brain
suffers from distortions originated by the standard 
\gls*{epi} sequences used in acquisition. These artifacts
are present at interfaces between tissues, due to the inhomogeneity
of the field derived from the discontinuity of magnetic
susceptibility. As a result, registered signal shows characteristic
geometrical deformations and signal destruction. These artifacts
suppose a significant pitfall that limits the success of tractography 
algorithms.

Several retrospective correction techniques are readily
available. In this work, we use a renowned digital phantom 
designed for evaluating tractography. We subject the phantom 
to a ``theoretically correct'' and plausible deformation that
resembles the artifact under investigation. We correct data back, 
with three standard methodologies (namely fieldmap-based, 
reversed encoding-based, and registration-based). 
Finally, we rank the methods based on their geometrical 
accuracy, the dropout compensation, and their impact on 
the resulting tractography.

\end{abstract}