### 2013-09-19 Publications/2013-ISBI14/05-conclusions.tex

\section{CONCLUSION}

This paper proposes an evaluation framework to
comprehensively analyze the impact of susceptibility
induced distortion on tractography from \gls*{dmri}
data. Inaccuracy on tractography leads to significant
errors in posterior connectivity analyses of the
whole-brain, specially in distortion affected regions.
We publicly release the framework and also contribute
with the evaluation of three widely-used available
correction methodologies. We conclude that the 
\gls*{reb} method resulted the first method in the 
proposed rankings (geometric correctness, signal loss
recovery and tractography impact), with the drawback
of demanding some extra \gls*{dmri} data.