### 2013-09-30 Publications/2013-ISBI14/05-conclusions.tex

\section{CONCLUSION}

This paper proposes an evaluation framework to
comprehensively analyze the impact of susceptibility
induced distortion on tractography from \gls*{dmri}
data. Inaccuracy on tractography produce errors
and increased variability in subsequent connectivity 
analyses of the whole-brain. Specially, in regions 
influenced by  significant distortion.
We publicly release the framework and also contribute
with the evaluation of the most widely-used and
readily-available correction methodologies. 
This work shows that \gls*{dmri} should be corrected
for susceptibility-induced artifacts prior to tractography
and connectivity analyses. In that sense, the \gls*{reb} 
method was the best classified
in two of the three rankings (geometric 
correctness and signal recovery), whereas \gls*{fmb}
achieved the lowest impact on the connectivity matrix
extracted from tractography.