### 2013-10-01 Publications/2013-ISBI14/05-conclusions.tex

\section{CONCLUSION}

This paper proposes an evaluation framework to
comprehensively analyze the impact of susceptibility
induced distortion on tractography from \gls*{dmri}
data. Inaccuracy on tractography produce errors
and increased variability in subsequent connectivity 
analyses of the whole-brain. Especially, in regions 
influenced by  significant distortion.
We publicly release the framework and also contribute
with the evaluation of the most widely-used and
readily-available correction methodologies. 
The reverse-encoding method (\gls*{reb}) achieved the
best results correcting for susceptibility-induced 
artifacts. We reported geometrical accuracy, signal
dropout recovery and connectivity matrix detection as
ranking indices.