### 2013-10-01 Publications/2013-ISBI14/01-abstract.tex
Final version

\begin{abstract}
Connectivity analysis on diffusion MRI data of the whole-brain
suffers from distortions caused by the standard 
echo-planar imaging acquisition strategies. %
%These artifacts are present at interfaces between tissues,
%due to the inhomogeneity of the field derived from the 
% discontinuity of magnetic susceptibility.
These images show characteristic geometrical deformations
and signal destruction that are an important pitfall
limiting the success of tractography algorithms.

Several retrospective correction techniques are readily
available. In this work, we use a digital phantom designed
for the evaluation of connectivity pipelines. We subject
the phantom to a ``theoretically correct'' and plausible
deformation that resembles the artifact under investigation.
We correct data back, with three standard methodologies
(namely fieldmap-based, reversed encoding-based, and
registration-based). Finally, we rank the methods based
on their geometrical accuracy, the dropout compensation,
and their impact on the resulting connectivity matrices.
\end{abstract}%
\begin{keywords}
susceptibility artifacts, diffusion MRI, tractography, connectivity.
\end{keywords}