### 2013-10-10 Publications/2013-ISBI14/04-results.tex
Added new results and figure

\section{RESULTS}
\label{sec:results}

All the modules described in the previous Section were
conveniently integrated in workflows using \emph{nipype}
\cite{gorgolewski_nipype:_2011}. The choice of this tool
grants the reproducibility of the experiments,
and the evaluation workflow is publicly released.
The results of the proposed experiments are summarized
in \autoref{fig:results} and \autoref{table:results}.
The remaining of this section provides extended descriptions
and interpretation of results.

\begin{figure*}[tpb]
   \centering
   \includegraphics[width=0.9\textwidth]{Fig02-Results}
   \caption{Visual comparison of correction methods results. 
   First row represents
   a coronal section of the \textit{b0} volume. In second row, the outcome
   of tractography after filtering tracks that did not connect two regions.
   Third row shows the associated connectivity matrices. First column shows
   the ground-truth. For the remaining steps, false positives have been marked
   in red color and false negatives in blue. Underlaying in gray color,
   the true positive (correctly detected) connections. }
   \label{fig:results}
\end{figure*}

\subsection{Geometrical correctness and signal recovery}

The three methods under study achieved good results in
terms of geometrical accuracy, as it is shown in the first
row of \autoref{fig:results}. The best numerical scores
were obtained with the \gls*{reb} method (\autoref{table:results}),
that achieved a weighted average overlap of $96.64\%$. Second
best were obtained with \gls*{fmb}. Geometrical correctness
is fundamental in connectivity analysis to spatially locate the
\glspl*{roi} that will define the nodes of the final connectivity
matrix. The clear difference of accuracy between \gls*{reb} and \gls*{fmb}
with respect to \gls*{t2b} infers the latter may not be
an appropriate method for susceptibility correction.

\input{041-results-table}

A second workflow investigated the similarity of the recovered
signal with respect to the original (undistorted).
In this second study, \gls*{reb} performed
better than the other two methods, as reported
in \autoref{table:results}. \gls*{reb} scored a $91.0\%$
similarity index for the \textit{b0} volume and an average $97.65\%$
for the remaining 31 \glspl*{dwi}. Again, the second
qualified was \gls*{fmb}, which achieved very close results
for the \glspl*{dwi} ($96.26\%$) but not as good for the \textit{b0}
volume. Visual inspection of the recovered data confirms the 
presented numerical results (\autoref{fig:results}, first row).

\input{042-results-tractography}

\subsection{Impact on tractography and connectivity}

Connectivity matrices derived from diffusion tensor imaging 
(using 32 directions \glspl*{dwi}) were completely
hindered by the high complexity of the fibers contained
in the phantom. Tractography discovered successfully only 4 out 
of 27 existing connections from the original (undistorted) data,
with more than 45 false connections.
With the 64-directions dataset (reconstructed with constrained 
spherical deconvolution) we found 23/27 connections,
but we still observed 40 false positives. Therefore, the 
experiments using the 32-directions phantom were discarded. The
results presented in this subsection only refer to the 
64-directions dataset.

Using the strongest distortion (maximum displacement of 7.60 mm),
the number of detected connections
remained the same (23/27), slightly increasing the number of
false positives to 42. The immediate conclusion is that, with highly
complex anatomies (crossing, fanning, etc.) and limited number of
ground-truth connections, connectivity matrices are more sensitive 
to reconstruction and tractography than to distortions.

For the sake of completeness, \autoref{table:results-tractography}
reports the characteristic features of the tractography results
and connectivity matrices, for the fieldmap that caused a maximum
displacement of 3.80 mm. Very similar results were obtained for
7.60 mm, a distortion which is at the limit of the real case. This
results, along with visual inspection (\autoref{fig:results}, second
and third rows) might point to \gls*{fmb} as the best correction method.

\subsection{Discussion}

Even though all the surveyed methods produced visually 
sound results, our study suggested that \gls*{reb} is the 
most accurate methodology for correction of susceptibility-induced
distortions in \gls*{dwi}. Moreover, \gls*{reb} also achieved
the best results in terms of connectivity, being able to recover
one of two lost connections after distortion. \Gls*{fmb} achieved 
very positive scores. Finally, the \gls*{t2b} method did not achieve
the necessary high-standards to recommend its use. We understand
that specific methods with anisotropic regularization that completely
restrict deformations to the phase-encoding direction would perform
significantly better compared to the standard method included
in this study.
%Finally, \gls*{t2b} confirmed that its correction
%can not be compared, even when using the intensity compensation
%based on the determinant of the deformation field.
%Although track counting has been used as a measurement of success and as a weight of connectivity 
%strength, the number of tracks produced by tractography strongly rely
%on the seeding performed and the smoothing caused by interpolation used
%within correction methodologies.

One relevant limitation of this work is that the \gls*{fmb} is
used both in distortion creation and correction, which could
bias the results. For practical reasons (distortion on the mask, 
signal dropout), the \gls*{fmb} can not achieve a perfect 
geometrical correction as might be
expected if the process was exactly inverse.
Nonetheless, \gls*{reb} performed better than \gls*{fmb}.
A second limitation is found on the impossibility of integrating the
point spread function-based methods for two reasons.
Firstly, the unmet need of a point spread function map consistent
with the synthetic phase-difference map. Secondly, to our knowledge,
there are no implementations of the method readily available.
In practice, this second limitation is narrowed down by the
fact that typical acquisition protocols usually include one or more
datasets among field mapping, reversed-encoding \textit{b0} or T2,
but they do not include the point-spread function mapping.

Future extensions of this work will evaluate
different realizations of the synthetic phase-difference map,
better characterizing the phenomenon. Moreover, the
framework can be enhanced for studying the impact of 
other artifacts as subject's motion or eddy 
currents-derived distortions.