### 2013-10-11 Publications/2013-ISBI14/05-conclusions.tex

\section{CONCLUSION}

This paper proposes an evaluation framework to
comprehensively analyze the impact of susceptibility
induced distortion on tractography from \gls*{dmri}
data. Inaccuracy on tractography produce errors
and increased variability in subsequent connectivity 
analyses of the whole-brain. This effect is suspected
to be larger in regions influenced by large distortions.
Even though, this study found out that the connectivity
matrix is more largely affected by the tractography method.
We publicly release the framework and also contribute
with the evaluation of the most widely-used and
readily-available correction methodologies. 
The reverse-encoding method (\gls*{reb}) achieved the
best results correcting for susceptibility-induced 
artifacts. We also reported visual evaluation of 
tractography and quantitative results for the 
extracted connectivity matrices.