### 2013-10-11 Publications/2013-ISBI14/05-conclusions.tex
Final version submitted before new deadline (Oct 11)

\section{CONCLUSION}

This paper proposes an evaluation framework to
comprehensively analyze the impact of susceptibility
induced distortion on connectivity analyses from \gls*{dmri}
data. Inaccuracy on tractography produce errors
and increased variability in the extracted connectivity 
matrices of the whole-brain. This effect is suspected
to influence those regions with larger distortions.
However, this study found out that the connectivity
matrix is more largely affected by the tractography method.
We publicly release the framework and also contribute
with the evaluation of the most widely-used and
readily-available correction methodologies. 
The reverse encoding-based method achieved the
best results correcting for susceptibility-induced 
artifacts. We also reported visual evaluation of 
tractography and quantitative results for the 
extracted connectivity matrices.