### 2013-12-10 Publications/2013-ISBI14/02-introduction.tex
Final proposal and response letter draft

\section{INTRODUCTION}
\label{sec:intro}
In-vivo whole-brain connectivity analysis has been a research topic of high
  interest for the last 5 yr.
\Gls*{dmri} can be used to probe the orientation of fiber bundles within the
  brain, and it is generally acquired with an \gls*{epi} sequence.
After signal reconstruction, tractography algorithms draw a map of the sampled
  structures.
These maps can represent the actual trajectories of fiber bundles
  (deterministic tractography) or pixel-wise probability of connection
  to a certain origin (probabilistic tractography).
Finally, the information about these connections is collected into a network
  matrix that is subjected to the ``connectome analysis''.

Among all the difficulties that such a complex workflow raises, % \cite{jones_twenty-five_2010},
  we will address here the susceptibility-derived artifacts,
  for which \Gls*{epi} schemes are especially sensitive.
Magnetic susceptibility disturbs the magnetic field close to tissue interfaces.
This inhomogeneity of the field translates into a highly distorted anatomy
  and a significant signal destruction in certain regions of the brain 
  (e.g. the orbitofrontal lobe, for the proximity of the air surrounding sinuses).
This artifact has been well described, generally within the context of functional MRI,
  which also uses \gls*{epi}.

Various approaches have been proposed to correct for this distortion.
\Gls*{fmb} methods \cite{jezzard_correction_1995} rely on one extra
  acquisition (\emph{field mapping}), that probes the inhomogeneity of the
  B0 field.
A second theory-based breed of methodologies acquire a map of the \acrlong*{psf}
  of the \gls*{epi} readouts to correct for the artifact \cite{robson_measurement_1997}.
Another approach referred to as \gls*{reb}, acquires an extra \gls*{epi} volume in the
  orthogonal or reversed phase encoding that can be combined to remove the geometric
  distortions \cite{cordes_geometric_2000,chiou_simple_2000}.
The last set of methodologies uses an extra T2-weighted image as anatomical reference
  to seek for the deformation map through nonlinear registration 
  \cite{kybic_unwarping_2000,studholme_accurate_2000}.
These \gls*{t2b} methods usually map the T2w image to the \emph{baseline} volume or
  \textit{b0} of \gls*{dmri}, as the latter is a T2 weighted image as well.
More recent works report extensions or combinations of existing techniques 
  \cite{andersson_how_2003,zaitsev_point_2004,%
  holland_efficient_2010,andersson_comprehensive_2012}.

Even though the aforementioned techniques for distortion correction have been studied
  \cite{zeng_image_2002,wu_comparison_2008}, the lack of a gold-standard limits benchmarking
  strategies.
Recently, Irfanoglu et al. \cite{irfanoglu_effects_2012} raised the question of distortion-derived
  impact on tractography.
In this work, we propose an evaluation framework using a digital phantom designed for connectivity
  assessment.
This framework enabled us to compare several correction techniques and characterize their geometrical
  accuracy and the dropout compensation.
Finally, we report their impact on subsequent tractography and connectivity.