### 2013-12-10 Publications/2013-ISBI14/04-results.tex

\section{RESULTS}
\label{sec:results}
All the modules described in the previous section were
  conveniently integrated in workflows using \emph{nipype} % \cite{gorgolewski_nipype:_2011}
  (\url{nipy.sourceforge.net/nipype}).
The choice of this tool grants the reproducibility of the experiments,
  and the evaluation workflow is publicly released.
The results of this study are summarized in \autoref{fig:results} and
  \autoref{table:results}.
The remaining of this section provides extended descriptions and interpretation of these
  findings.
\begin{figure*}[tp]
   \centering
   \includegraphics[width=0.95\linewidth]{Fig02-Results}
   \caption{Visual assessment of the correction methods.}
   \label{fig:results}
\end{figure*}

\subsection{Geometrical correctness and signal recovery}\label{sec:results_geometry}
A summary of quantitative results, computed for the \gls*{dti} dataset with max. shift 3.80~mm,
  is provided in \autoref{table:results}.
The best scores were obtained with the \gls*{reb} method, followed by \gls*{fmb}.
The clear difference of accuracy between \gls*{reb} and \gls*{fmb} with respect to \gls*{t2b}
  infers the latter may not be an appropriate method for susceptibility correction.
The described performance was constant for \gls*{fmb} and \gls*{reb} along the different magnitudes 
  of distortion evaluated (max. shift 3.80-7.60~mm).
\input{041-results-table}
\input{042-results-tractography}
A second workflow investigated the similarity of the recovered signal with respect to the original
  (undistorted).
In this second study, \gls*{reb} performed better than the other two methods, as reported in
  \autoref{table:results}.
\Gls*{reb} scored a 91.0\% similarity index for the \textit{b0} volume and an average 97.65\%
  for the remaining \glspl*{dwi}.
Again, the second qualified was \gls*{fmb}, which achieved very close results for the \glspl*{dwi}
  (96.26\%) but not as good for the \textit{b0} volume.
Visual inspection of the recovered data confirms the presented quantitative results 
  (\autoref{fig:results}).

\subsection{Impact on tractography and connectivity}
Connectivity matrices derived from the \gls*{dti} dataset were completely hindered by the high complexity
  of the fibers contained in the phantom.
Tractography discovered successfully only 4 out of 27 existing connections from the original (undistorted) data,
  with more than 45 false connections.
With the \gls*{hardi} dataset we found 23/27 connections, but we still observed 40 false positives.
Therefore, the experiments using the \gls*{dti} phantom were discarded.
The results presented in this subsection only refer to the \gls*{hardi} dataset.

Using the strongest distortion (max. shift 7.60~mm), the number of detected connections
  remained the same (23/27), slightly increasing the number of false positives to 42.
The immediate conclusion is that, with highly complex anatomies (crossing, fanning, etc.)
  and limited number of ground-truth connections, connectivity matrices are more sensitive
  to reconstruction and tractography than to distortions.

For the sake of completeness, \autoref{table:results-tractography} reports the characteristic
  features of the tractography results and connectivity matrices, for the fieldmap that caused a max.
  shift of 3.80~mm.
A very similar outcome was obtained for 7.60~mm.
These results, along with visual inspection (\autoref{fig:results}), might point to \gls*{fmb} as
  the best correction method.

\subsection{Discussion}\label{sec:discussion}
Even though all the surveyed methods produced visually sound results, our study suggested that
  \gls*{reb} is the most accurate method in terms of geometrical agreement and signal dropout recovery.
The \gls*{t2b} method did not achieve the necessary high-standards to recommend its use.
Nonetheless, we understand that specific methods with anisotropic regularization that completely
  restrict deformations to the phase-encoding direction would perform significantly better than
  the standard method presented.
Geometrical correctness of \gls*{dwi} data is fundamental in connectivity analysis to spatially
  locate the regions which will define the nodes of the final connectivity matrix.

Regarding tractography, this study revealed that signal reconstruction and tractography algorithms
  masked the impact of susceptibility distortion on the final connectivity matrices.
Due to this effect, experiments performed on the \gls*{dti} phantom were discarded for comparison.
With the \gls*{hardi} dataset, the extracted connectivity matrix slightly changed with distortion.
Quantitative differences reported in \autoref{table:results-tractography} could be more related to
  the smoothing derived from interpolation implemented by each method.
Visual results might suggest that \gls*{fmb} achieved better results.

Although we found connectivity matrices rather invariant with respect to distortion, they are suspected
  to be significantly impacted by the susceptibility artifact on real data \cite{irfanoglu_effects_2012}.
This hypothesis points to the need of more appropriate phantoms with denser connectivity matrices.
State of the art phantoms for tractography usually present a discrete set of simulated fiber bundles that
  translate in very sparse fiber-endings regions and connectivity matrix.
In the real case, the surface limiting the tracks is densely covered by the regions mapped from the anatomical
  dataset, what leads to larger sensitivity with respect to deformations.

A possible limitation of this work is that \gls*{fmb} is used for both synthesis and correction of the
  distortion.
However, practical reasons (i.e. noise, signal dropout) impede perfect geometrical correction with \gls*{fmb}.
Moreover, \gls*{reb} performed more accurately than \gls*{fmb}.

Future extensions of this work will include a refined digital phantom.
Additional lines for this work will evaluate different realizations of the synthetic phase-difference map,
  to better characterize the phenomenon.
Also, the framework can be enhanced for studying the impact of other artifacts as subject's motion or eddy
  currents-derived distortions.