### 2014-04-11 Publications/2014-TMI/02-abstract.tex
New paper branch

In whole-brain connectivity analysis of diffusion MRI (dMRI) data,
an accurate delineation of the white-matter and grey-matter 
interfaces is required. While high-standard segmentation is readily 
available for structural MRI, dMRI typically present significant cerebrospinal
fluid contamination effects, mainly derived from its typical low resolution,
and severe geometrical distortions. We propose a segmentation-registration 
variational framework that exploits the detailed anatomy extracted from 
structural MRI as shape-prior. We use an ``active contours without edges''-like 
model to search for a deformation field that optimally maps the shape prior 
on multivariate dMRI-derived data in diffusion space, registering structural
and diffusion coordinate spaces and implicitly segmenting dMRI data. The approach
proven the intrinsic coupling of segmentation and distortion correction and
we evaluate its results on a digital simulated phantom and real datasets. 
Therefore, precise and consistent cortical parcellation on dMRI is 
straightforward by projection from T1 space, avoiding additional registration,
segmentation and/or surface matching steps.