### 2014-04-22 Publications/2013-ISBI14/poster/00-main.tex
Starting with poster...

% @Author: oesteban
% @Date:   2014-04-22 08:55:40
% @Last Modified by:   oesteban
% @Last Modified time: 2014-04-22 12:22:50

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}
\setbeamertemplate{bibliography item}[text]

\let\Tiny=\tiny
\providecommand\thispdfpagelabel[1]{}
\mode<presentation>{\usetheme{Blue}}
\usepackage[english]{babel}

\usepackage[T1]{fontenc}
\usepackage{textcomp}

\usepackage[natbib=true, style=numeric-comp,backend=biber]{biblatex}

\usepackage{amsmath,amsthm, amssymb, latexsym}
\usepackage{lmodern}
\usefonttheme[onlymath]{serif}
%\boldmath
\usepackage[orientation=portrait,size=a0,scale=1.4]{beamerposter}

\usepackage{graphicx,epstopdf}
\DeclareGraphicsExtensions{.pdf,.eps}
\graphicspath{{../figures/} {logos/}}

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
\newcommand{\sups}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subs}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\pphantom}{\textcolor{ta2skyblue}} % phantom introduces a vertical space in p formatted table columns??!!
\usepackage[scaled=.7]{beramono}

\bibliography{99-references}
\appto\bibfont{\hskip1cm\tiny}

\DeclareSymbolFont{cmlargesymbols}{OMX}{cmex}{m}{n}
\let\sumop\relax
\DeclareMathSymbol{\sumop}{\mathop}{cmlargesymbols}{"50}

\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\huge Simulation-based evaluation of susceptibility distortion correction methods in dMRI}
\author[Esteban et al.]{O. Esteban$^{1,2}$\and%
A. Daducci$^{2}$\and%
E. Caruyer$^{3}$\and%
K. O'Brien$^{4}$\and%
MJ. Ledesma-Carbayo$^{1}$\and%
M. Bach-Cuadra$^{5,2}$\and%
A. Santos$^{1}$}%}
\institute{
$^{1}$~\includegraphics[height=2cm]{logos/BIT.pdf}~\includegraphics[height=2cm]{logos/UPM.pdf}\hfill%
$^{2}$~\includegraphics[height=2cm]{logos/LTS5.pdf}~\includegraphics[height=2cm]{logos/EPFL.pdf}\hfill%
$^{3}$~\includegraphics[height=2cm]{logos/SBIA.pdf}~\includegraphics[height=2cm]{logos/Penn.pdf}\hfill%
$^{4}$~\includegraphics[height=2cm]{logos/CIBM.pdf}~\includegraphics[height=2cm]{logos/UNIGE.pdf}\hfill%
$^{5}$~\includegraphics[height=2cm]{logos/CHUV.pdf}~\includegraphics[height=2cm]{logos/UNIL.pdf}%
}
\date[May 5th, 2012]{May 5th, 2012}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newlength{\columnheight}
\setlength{\columnheight}{105cm}

\begin{document}

\begin{frame}
  \begin{columns}
    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            %\input{motivation}
%            \vfill
            %\input{methods}
%            \vfill
            %\input{conclusions}
            %\vfill
          }
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content
	    %\input{dataset}
        %\input{results}
        %\input{acknowledgements}
          }
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column
  \end{columns}
\end{frame}

\end{document}