### 2014-10-20 Publications/2013-TMI/03-introduction.tex
update TMI paper

% -*- root: 00-main.tex -*-
\section{Introduction}
\label{sec:introduction}

\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
  of \gls{mri} techniques \citep{sundgren_diffusion_2004} which
  accounts for a growing interest in its application to structural
  connectivity analyses of the brain at the macro-scale
  \citep{craddock_imaging_2013}.
This emerging field exploits \gls{dwi} data to derive the local axonal structure at
  each imaging voxel \citep{basser_microstructural_1996} and estimate a whole-brain
  mapping of fiber tracts \citep{hagmann_mr_2010,berg_using_2009} trajectories
  reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the
  \emph{connectome} \citep{hagmann_from_2005,sporns_human_2005}.
The connectome analysis is a promising tool for neuroscience \citep{morgan_why_2013}
  and for the clinical application \citep{griffa_structural_2013} as well.

% REMOVE paragraph (Dom, 1st comment)
%Early \gls{dwi} research mainly focused on the improvements of imaging
%methodologies by a better understanding of the diffusion effect, and improving
%image reconstruction methodologies.
%Currently, the connectome extraction
%and analysis relies on a large chain of sophisticated computational techniques
%including acquisition, reconstruction, modeling and model fitting, image segmentation
%and registration, fibre tracking, connectivity mapping, visualization, statistical
%group studies, and inference \citep{daducci_connectome_2012}.
%This growing complexity
%has given rise to challenging issues towards reliable structural information about
%the neuronal tracts presented in, and also difficulties on the statistical analysis
%\citep{bullmore_complex_2009,meskaldji_comparing_2013}.
%To our knowledge, studies
%regarding the impact of the different sources of inaccuracy and variability along
%the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
%cheng_characteristics_2012,jones_white_2013} are scarce.

\subsection{\Gls{dwi} data overview}
\label{sec:dwi_overview}
\gls{dwi} data are usually acquired with \gls{epi} schemes as they allow for very
  fast acquisition of large sets of slices each acquired after a single excitation
  \citep{mukherjee_diffusion_2008,soares_hitchhikers_2013}.
\Gls{epi} volumes are collected in sequence and each one represents one probed
  diffusion direction.
%This first acquisition is called \emph{raw} data or \emph{diffusion
%signal}.
%One or more volumes without directional encoding (so-called \emph{baseline},
%$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
%reference, at the least one is acquired at the beginning of the
%\gls{epi} dataset.
%These B0 are acquired without direction gradient,
%showing a T2-like contrast.

\Gls{epi} are known to suffer from geometrical distortions and artifacts
  due to, mainly, three sources: the subject motion in between acquiring
  different sampling directions, the induced \emph{eddy currents} on the scanner
  coils due to the gradient switching \citep{brien_3d_2013}, and, finally,
  distortions caused by the magnetic susceptibility inhomogeneity
  present at tissue interfaces.
\emph{Susceptibility distortions} happen along the phase-encoding direction,
  and are most appreciable in the frontal part of the brain due to the air/tissue
  interface located around the sinuses.
Two implications are associated to this artifact: the signal loss caused on
  highly distorted regions, and a significant added difficulty in
  registering with structural images (e.g.
T1-weighted).

After correcting for artifacts, the difficulty of performing analyses directly
  on the \emph{directionally dependent} raw data is not affordable.
Thus, \gls{dwi} data is reconstructed using one of a numerous set of readily
  available algorithms.
The outcome is a voxel-wise model of diffusion shape (\citep{daducci_quantitative_2014}
  contains a comprehensive review of state-of-art methodologies).
On top of the chosen model, some scalar maps describing tissue features are generally
  computed (e.g.
\gls*{fa}, and \gls*{adc}).
%For instance, the most popular \gls{dwi}-derived scalar map
%is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
%Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
%which reflects the intensity of the diffusion effect.
%Many other
%directionally invariant measurements have been defined.
%For a comprehensive study,
%we refer to \citep{ennis_orthogonal_2006}.

Alongside distortions and direction dependency, \gls{dwi} data is strongly affected
  by \glspl{pve} \citep{alexander_analysis_2001} that appears when several different
  tissues are present in the same imaging unit, producing an averaged intensity.
The effect is directly related to the resolution, typically around
  $2.0\times2.0\times2.0mm^3$ for \gls{dwi}.

% the low resolution achievable with \gls{dwi}
%(typically around ).
%An additional complication specific to
%\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
%that is a particular \gls{pve} in which the signal sensed inside the affected voxel is
%linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
%respect \gls{csf}.

\subsection{Correction for susceptibility distortions}
\label{sec:distortion}

One approach to correcting the susceptibility distortions was proposed with the
  earliest \gls{epi} sequences used in functional \gls{mri}, and relies on the
  acquisition of extra \gls{mri} data.
Generally, a \gls{gre} is used to obtain a map of magnitude and phase of the actual
  magnetic field inside the scanner.
Based on this \emph{fieldmap} and the theory underlying the distortion,
  \citep{jezzard_correction_1995} proposed a correction methodology.
A number of forked and improved methodologies have been developed to correct
  for the susceptibility distortion and received the generic name of
  ``fieldmap correction'' techniques \cite{hsu_correction_2009,reber_correction_1998}.

Some other retrospective methodologies do not make use of the fieldmaps, as explicitly
  modeling the distortion \citep{andersson_modeling_2001}, registering with (anatomically
  correct) T2-weighted \gls{mri} \citep{kybic_unwarping_2000,studholme_accurate_2000,
  li_accounting_2007,tao_variational_2009}, or acquiring an extra B0 image with reversed
  phase encoding direction \citep{andersson_how_2003,holland_efficient_2010}.

A comparison of the diverse correction techniques is found in \citep{wu_comparison_2008}.
This study claims that fieldmap correction methodologies are not entirely accurate
  and reliable, even though the method is correct in principle.
The conclusion was later confirmed by \citep{tao_variational_2009}.
Additional concerns regarding fieldmap correction are the requirement of an extra acquisition
  (that is not always met or it is impractical), or the accuracy of the measured fieldmap that
  is sensitive to various effects (such us respiration, blood flow, etc.).
All these factors have turned susceptibility distortion in \gls{epi} sequences an active field of
  research for the last 10 years.

\subsection{\Gls{dwi} segmentation}
\label{sec:dwi_segmentation}

A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
  is required with sub-pixel resolution, because whereas tractography algorithms are
  very good at estimating the location of bundles in deep \gls{wm}, they are not
  god (yet) at identifying where the tracts project into the \gls{gm}
  (limitations on the termination criteria, \citep{jbabdi_tractography_2011,
  craddock_imaging_2013}).
Thus, a \gls{wm} segmentation is required by most of the available tractography
algorithms to \emph{filter} the resulting tractogram.
The \gls{gm}-\gls{wm} interface is necessary to locate the starting and ending points
  of the detected fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning
  spurious and discontinued fiber bundles.

A number of methodologies have been proposed for \gls{dwi} segmentation, ranging
  from intensity thresholding to atlas-based segmentation.
The first approach is performed simply thresholding the \gls{fa} map.
Although this methodology was popular among the pioneer tractography studies,
  they were generally limited to certain regions or significant fiber tracts, and thus,
  it cannot be applied in whole-brain tractography.
Early approaches to \gls{dwi} segmentation
  include level set formulations using scalar maps of direction invariants derived
  from the tensor model \citep{zhukov_level_2003}, directly on the diffusion raw data
  \citep{rousson_level_2004}, or finding alternative diffusion representations
  \citep{jonasson_representing_2007}.
Even though this latter case was restricted to the extraction of
  the corpus callosum from a real dataset, the density of the components of the diffusion tensor
  are approximated by multivariate Gaussians for first.
Iterative clustering performed on the B0 volumes of \gls{dwi} data was proposed by
  \citep{hadjiprocopis_unbiased_2005}.
Later studies investigated the application of probabilistic frameworks combining mixtures of
  gaussians, \gls{mrf} and labeling fusion techniques \citep{liu_brain_2007} using as features
  widely-used \gls{dwi}-derived scalar maps as \gls{fa} or \gls{md}.
A similar framework combining co-registered structural information (T1 weighted) with
  \emph{independent orthogonal invariants} derived from the \gls{dwi} tensor model was proposed
  by \citep{awate_multivariate_2008}.
Some proposals suggest the use of the raw diffusion data (directionally dependent) to avoid fitting
  a certain model \citep{lu_segmentation_2008}.
In \citep{han_experimental_2009}, graph-cuts voxel-based techniques are proposed using the most
  common diffusion tensor derived features.
Further developments of the probabilistic approach have been proposed adding more scalar maps
  as features and a more detailed treatment of \gls{pve} \citep{kumazawa_partial_2010}.

%A number of methods have been proposed using features not directly derived from \gls{dwi} data.
%Segmentation obtained by co-registering structural T1-weighted images will be covered in
%\autoref{sec:registration}.

None of the presented methods have claimed for definite results, mainly due to the lack of
  a \emph{gold-standard} evaluation methodology.
Most of them are tested only on certain
  regions, or do not provide sub-pixel resolution results
  \citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
  han_experimental_2009}.
Generally, results obtained with high resolution atlas co-registration
  (\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi}
  segmentation.
However, recently \citep{kumazawa_improvement_2013} proposed to include this
  task within reconstruction algorithms.
Finally, some \emph{golden}-standard evaluation frameworks for task-based validation have been
  proposed for the segmentation results (e.g. lesion detection in visceral organs
  \citep{jha_taskbased_2012}).
\todo[inline]{reference to a summary table}

\subsection{Structural information co-registration}
\label{sec:registration}

The last addressed task in the image processing stage of the studied
connectivity analysis pipelines is the structural information co-registration.
The need for this registration step is mainly raised when defining the
nodes of the connectivity matrix.
This definition step is named
\textit{cortical parcellation} as it imposes the regions that will be
considered to cluster the fiber bundles reaching the \gls{wm}-\gls{gm}
surface.
The common procedure to accomplish this clustering is using
a predefined parcellation in a high-resolution structural T1-weighted
atlas (e.g.
\citep{greve_accurate_2009}).
The second goal that has
justified the applicability of structural
images co-registration is susceptibility distortion correction
as presented in \autoref{sec:distortion}.
In any case, susceptibility
distortion hinders a rigid-registration solution to this problem.
Additionally, when applying non-linear intensity-based
registration algorithms, other difficulties have to be addressed as
the significant \gls{pve} in the \gls{wm}-\gls{gm} layer, or the
inherent inadequacy of the B0 contrast (the only \textit{directionally
independent} existing in the raw \gls{dwi} data) due to the almost null
difference in intensity between \gls{wm} and \gls{gm} pixels.
Early
registration methods appeared targeting the distortion correction.
They quickly standardized the choice of T2 as anatomically correct source
to be registered against the B0 in \gls{dwi}.
B0 images have a very similar
contrast to T2 weighted due to the parallelism of acquisition sequences.
However, parcellation is defined in T1 weighted space, adding an additional
registration step is generally overlooked by existing literature.
Even though this
processing stage has a very low complexity compared to the remaining tasks of
the pipeline process, it is one additional source of inconsistency
and unreliability.

One of the first proposals is \citep{kybic_unwarping_2000}, where the deformation
field is modeled with B-Splines, the cost function is least squares and optimized
in a multi-resolution gradient descent strategy.
Their method is evaluated in
both synthetic and real 2D images.
Similarly,
\citep{studholme_accurate_2000} proposed a spline-based deformation but including
a weighting factor proportional to the Jacobian of the transform to correct the
intensity of the undistorted data.
They also use the $\log$ of the signal to
enhance the low-signal regions, and optimize with a gradient descent algorithm
the mutual information of the mapping.
This framework set the basis for the following
works.
For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
providing quantitative comparisons with fieldmap correction methods.
Recent approaches
take into account the signal loss due to dephasing \citep{li_accounting_2007}, or
introduce more complex variational frameworks \citep{tao_variational_2009}.

The significant benefits of exploiting the anatomical \gls{mri} when
segmenting the \gls{dwi} data have been demonstrated \cite{zoellei_improved_2010},
justifying the use of the shape prior information.
To our knowledge, there is
no study simultaneously taking advantage from segmentation or distortion correction
tasks to be applied to the co-registration problem.

\subsection{Proposed model}
\label{sec:contributions}

All the available \gls{dwi}-derived models and scalar maps
are very sensitive to distortions, \gls{pve}, noise and axonal dispersion
\citep{jbabdi_tractography_2011}, what yields numerous challenges in the
processing tasks (artifact correction, segmentation, registration and
parcellation).
Finally, as a result a vast number of challenges and pitfalls
emerge regarding accuracy (correctness) and precision (reproducibility) of
tractography-based analyses \citep{jones_white_2013} and connectome
analyses.

Moreover, the problems of precise segmentation in \gls{dwi}-space and the
spatial mapping between these contours and the corresponding surfaces in
anatomical images bear significant redundancy.
Once the spatial relationship
between \gls{t1} and \gls{dwi} space is established, the contours which are
readily available in \gls{t1} space can simply be projected on to the
\gls{dwi} data.
Conversely, if a precise delineation in \gls{dwi} space
was achieved, the spatial mapping with \gls{t1}-space could be derived
from one-to-one correspondences on the contours.
However, neither segmentation
nor registration can be performed flawlessly, if considered independently.

% Paragraph taken from first subsection to here
Here, we shall address three tasks included within  the image processing stage in
a unified approach: brain tissue segmentation in diffusion space
(\autoref{sec:dwi_segmentation}),
correction of geometrical distortions (\autoref{sec:distortion}), and structural
image registration to diffusion coordinate space (\autoref{sec:registration}).
These tasks are generally
solved independently, or combined in pairs.
However, there exist fundamental
coupling relationships that can be exploited to obtain a simultaneous solution to
the three problems.
This joint approach satisfactory impacts the downstream
outcomes of the whole pipeline with the increase of the internal consistency.

In this paper we propose a novel registration framework to simultaneously
solving the segmentation, distortion and cortical parcellation challenges,
by exploiting as strong shape-prior the detailed morphology extracted
from high-resolution and anatomically correct \gls{mri}.
Indeed, hereafter
we assume this segmentation problem in anatomical images is reliably and
accurately solved with readily available tools (e.g.
\citep{fischl_freesurfer_2012}).
After global alignment with \gls{t1} using existing approaches, the remaining
spatial mismatch between anatomical and diffusion space is due to susceptibility
distortions.
Finally, we need to establish precise spatial correspondence between the
surfaces in both spaces, including the tangential direction for parcellation.
Therefore, we can reduce the problem to finding the differences of spatial
distortion in between anatomical and \gls{dwi} space.
We thus reformulate the segmentation problem as an inverse problem, where we
seek for an underlying deformation field (the distortion) mapping
from the structural space into the diffusion space, such that the structural
contours segment optimally the \gls{dwi} data.
In the process, the one-to-one
correspondence between the contours in both spaces is guaranteed, and projection
of parcellisation to \gls{dwi} space is implicit and consistent.

We test our proposed joint segmentation-registration model on two different
synthetic examples.
The first example is a scalar sulcus model, where the
\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
segmented correctly thanks to the shape prior and its coupling with the inner,
\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
The second case deals with more realistic \gls{dwi} data stemming from
phantom simulations of a simplistic brain data.
Again, we show that the
proposed model successfully segments the \gls{dwi} data based on two derived
scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
of the dense distortion field.

The rest of this paper is organized as follows.
First, in \autoref{sec:methods}
we introduce our proposed model for joint multivariate segmentation-registration.
Then we provide a more detailed description of the data and experimental setup in
\autoref{sec:experiments}.
We present results in \autoref{sec:results} and conclude
in \autoref{sec:conclusion}.