### 2014-10-20 Publications/2013-TMI/02-abstract.tex
updating intro

% -*- root: 00-main.tex -*-
Integrating information from different images within- and between-subjects
  is an ubiquitous necessity required in current neuroimage analyses.
Many algorithms have been proposed to register image pairs from the same
  subject and between subjects to resolve the problem.
Both images are required to present comparable features on top of which
  an objective function is built to evaluate fitness of the solution.
However, oftentimes one of the images is not available or the objective
  function is not adequate or numerically tractable.

We propose a surface-to-volume variational solution to the registration problem,
  that exploits the detailed anatomy extracted manually or from structural images.
We use an ``active contours without edges''-like model to search for a
  deformation field that optimally maps the surface prior onto the data
  in the target space, implicitly segmenting with sub-pixel resolution the
  volume.
We demonstrate the usefulness of the approach on a digital phantom, and on
  10 real datasets randomly selected from the Human Connectome Project.