### 2014-10-21 Publications/2013-TMI/05-experiments.tex

% -*- root: 00-main.tex -*-
\section{Data and experiments}
\label{sec:experiments}

\subsubsection{Simulated digital phantom} %
A description goes here: surfaces extraction, synthetic susceptibility artifact generation.

\subsubsection {Real \gls{mri} data} %

We used a standard automated method
available in \emph{FreeSurfer} \citep{fischl_freesurfer_2012} to obtain the
cortical gray/white boundary from the T1 scan \citep{greve_accurate_2009}.
Parcellations used to evaluate the repeatability of the method are also
obtained from \emph{FreeSurfer}.