### 2014-10-21 Publications/2013-TMI/03-introduction.tex
updated all

% -*- root: 00-main.tex -*-
\section{Introduction}
\label{sec:introduction}

\IEEEPARstart{N}{onlinear} registration of images is a challenging task ubiquitous
  in an endless number of image analysis applications.
The problem is solved finding a spatial mapping $U$ that aligns the
  information given in two different coordinate systems (generally called
  \emph{reference}, $R$, and \emph{moving}, $M$):
  \begin{align}
  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
  \label{eq:transform}
  \end{align}
  where $\vec{r}$ denotes a position in the reference domain and $\vec{r}'$
  its corresponding location in the target, and $n$ the dimension of domains
  ($n=3$ for 3D images).
Finally, $\vec{u} = u(\vec{r})$ is the function that models the displacement
  of every point in the reference domain.

The wide range of applications has derived in an intensive generation of diverse
  registration methods.
In a recent and comprehensive survey by \citeauthor{sotiras_deformable_2013}
  \citep{sotiras_deformable_2013}, registration methods are classified by
  matching criteria, deformation model, and optimization method.
The most widespread matching criteria are probably the so-called \emph{iconic
  methods} in \citep{sotiras_deformable_2013}.
However they are not appropriate for all the registration problems
  \citep{greve_accurate_2009}.
They propose a \emph{hybrid method} \citep{sotiras_deformable_2013} for the
  intra-subject registration of \gls*{t1} \gls*{mri} and \gls*{dmri} images
  of the brain using a linear transformation model.
As in \citep{greve_accurate_2009}, we exploit the precise structures that
  can be readily extracted with widely-used tools at high resolution
  and map this prior information to low resolution data,
  in which the scale of the imaged structures is in the range of the image
  resolution or above.
Our approach differs with \citep{greve_accurate_2009} in two fundamental
  choices: 1) We use a nonlinear deformation model that enables the
  correction of the susceptibility-derived distortions
  \citep{jezzard_correction_1995} that \gls*{dmri} data typically present.
And 2) our matching criteria does not necessarily need for clear intensity
  steps at the tissue interfaces where surfaces must be fitted.

\subsection{Related work}
\label{sec:methods_background}
Registration-segmentation methods are typically built including a deformation
  model to guide the evolution of a segmentation algorithm.
Even though some Bayesian approaches have been proposed to solve the joint problem
  \citep{wyatt_map_2003}, segmentation is usually performed using an evolving
  active contour \citep{chan_active_2001} that normally includes a shape prior
  term to the energy functional \citep{chen_using_2002,bresson_variational_2006,
  chan_level_2005,cremers_kernel_2006,gastaud_combining_2004}.
These methods generally have an explicit description of the expected relative boundary
  locations of the object to be delineated, and some, even model the statistical deviations
  from this average shape.

An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
  and reference images with an affine transformation supporting the coordinates mapping
  (the so-called ``2 \glspl*{pde} approach'').
\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
  based on the level-sets framework and using only one \gls*{pde}.
The principal application of these techniques was the image registration between timesteps in
  a time-series or between different images \citep{paragios_level_2003}.
\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
  using a free deformation field model.
A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
  functional in joint registration and segmentation, thorugh the propagation of the deformation
  ield from the contours to the whole image definition.
Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
Even though the historical evolution of joint segmentation and registration procedures is strongly
  linked to the level sets approximation, our work is based in derivations of the \gls{acwe}
  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.

Finally, a derivation of the latter group is composed by atlas-based segmentation methods
  \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,pohl_bayesian_2006,
  wang_joint_2006}, where the prior imposes consistent voxel-based classification of contiguous
  regions.
A comprehensive summary of the methodologies derived from the \gls{acwe} model, with special attention
  to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.

%{\color{red} {The main diverging points with respect to
%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
%but we replace the Gaussian filtering by other constraints
%studied in \citep{nagel_investigation_1986} to better the problem;
%3) optimization is applied in the spectral
%domain, observing anisotropic and inhomogeneous mappings along each direction.
%With respect \citep{guyader_combined_2011}, the main differences are
%the distance function, and the spectral solution to the optimization updates,
%as we shall cover in \autoref{sec:numerical_implementation}.}}

\subsection{Susceptibility-derived distortions}

\subsection{Our framework}
\label{sec:our_framework}

%In this paper we formulate the joint registration-segmentation
%problem as follows.
%such that the known contours in anatomical space $T$ optimally segment
%the diffusion space $D$.
%
%
%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
%we alternatively use \emph{shape-gradients}
%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.

%
%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
%  its associated feature vector in the following.
%
%In our application, these surfaces are precise tissue interfaces of interest extracted
%  from a high-resolution, anatomically correct reference volume using
%  the well-established
%
%In this paper we propose a novel registration framework to simultaneously
%solving the segmentation, distortion and cortical parcellation challenges,
%by exploiting as strong shape-prior the detailed morphology extracted
%from high-resolution and anatomically correct \gls{mri}.
%Indeed, hereafter
%we assume this segmentation problem in anatomical images is reliably and
%accurately solved with readily available tools (e.g.
%\citep{fischl_freesurfer_2012}).
%After global alignment with \gls{t1} using existing approaches, the remaining
%spatial mismatch between anatomical and diffusion space is due to susceptibility
%distortions.
%Finally, we need to establish precise spatial correspondence between the
%surfaces in both spaces, including the tangential direction for parcellation.
%Therefore, we can reduce the problem to finding the differences of spatial
%distortion in between anatomical and \gls{dwi} space.
%We thus reformulate the segmentation problem as an inverse problem, where we
%seek for an underlying deformation field (the distortion) mapping
%from the structural space into the diffusion space, such that the structural
%contours segment optimally the \gls{dwi} data.
%In the process, the one-to-one
%correspondence between the contours in both spaces is guaranteed, and projection
%of parcellisation to \gls{dwi} space is implicit and consistent.
%
%We test our proposed joint segmentation-registration model on two different
%synthetic examples.
%The first example is a scalar sulcus model, where the
%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
%segmented correctly thanks to the shape prior and its coupling with the inner,
%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
%The second case deals with more realistic \gls{dwi} data stemming from
%phantom simulations of a simplistic brain data.
%Again, we show that the
%proposed model successfully segments the \gls{dwi} data based on two derived
%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
%of the dense distortion field.
%
%The rest of this paper is organized as follows.
%First, in \autoref{sec:methods}
%we introduce our proposed model for joint multivariate segmentation-registration.
%Then we provide a more detailed description of the data and experimental setup in
%\autoref{sec:experiments}.
%We present results in \autoref{sec:results} and conclude
%in \autoref{sec:conclusion}.
%
%
%The properties of the reconstructed tensors and derived scalar maps have
%been studied by \cite{ennis_orthogonal_2006}. Based on their
%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
%considered complementary features, and therefore we selected them for the
%energy model \eqref{eq:complete_energy} in driving the
%registration-segmentation process.
%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
%the \gls{md} is more related to the \emph{magnitude} of the process:
%
%\begin{align}
%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
%\end{align}
%where $\lambda_i$ are the eigenvalues of the diffusion tensor
%associated with the diffusion signal $S(\vec{q})$. There exist
%two main reasons to justify their choice.
%First, they are well-understood and standardized in clinical routine.
%Second, together they contain most of the information that is
%usually extracted from the \gls{dwi}-derived scalar maps
%\cite{ennis_orthogonal_2006}.
%