### 2014-10-21 Publications/2013-TMI/06-results.tex

% -*- root: 00-main.tex -*-
\section{Results and discussion}
\label{sec:results}

\subsection{Proof of concept}
\label{sec:results_phantom}

\subsection{Susceptibility distortion correction}
\label{sec:results_phantom}

\subsection{Discussion}
\label{sec:discussion}
Don't forget to mention other potential applications.

We implemented the methodology upon the widely used
Insight Registration and Segmentation
Toolkit~\footnote{\url{http://www.itk.org}} (ITK)
for its computational benefits, the standardized code, and
with the aim at making the procedure publicly available
when ready for sharing with the research community.