### 2014-10-21 Publications/2013-TMI/07-conclusion.tex

% -*- root: 00-main.tex -*-
\section{Conclusion}
\label{sec:conclusion}
%
We propose and make publicly available a novel methodology for
  simultaneous distortion correction and consistent parcellation
  of \gls*{dmri} images using a joint registration-segmentation
  framework.
From the Bayesian approach to image segmentation, we derive
  a cost function that models piecewise smooth regions with
  multivariate normal distributions.
This is, implicitly, an \acrlong*{acwe} model, in wich the
  contours evolve optimizing a B-Spline deformation
  model.
We demonstrate the aptness of the method with quantitative
  evaluations on digital phantoms and real \gls*{mri} data.