### 2014-10-21 Publications/2013-TMI/08-appendix-00-shapepriors.tex

% -*- root: 00-main.tex -*-

\section{Extended formulations}

\subsection{Shape priors}
\label{app:shape_priors}
Let $R(\vec{r})$ be an ``arbitrary'' function over the image domain,
and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
outer boundary as defined in \autoref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:
  \begin{align*}
  \frac{\partial}{\partial t} \int_\Omega R(\vec{r}') d\vec{r} &=
  \int_\Omega \frac{\partial}{\partial t}R(\vec{r}') d\vec{r} \notag\\
  &- \int_{\Gamma_{l,m}} R(\vec{r}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
  \end{align*}
  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
Assuming that the region descriptors $\Theta_l$ vary slowly enough, we can consider
  that $\frac{\partial}{\partial t} R(\vec{r}') = 0$ and thus:
  \begin{equation}
  \frac{\partial}{\partial t} \int_\Omega R(\vec{r}') d\vec{r} =
  - \int_{\Gamma_{l,m}} R(\vec{r}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r}
  \label{eq:shape_gradients}
  \end{equation}