### 2014-11-06 Publications/2013-TMI/02-abstract.tex
one more update to the paper

% -*- root: 00-main.tex -*-
Integrating information from different images within- and between-subjects
  is an ubiquitous necessity required in current neuroimage analyses.
Many algorithms have been proposed to register image pairs from the same
  subject and between subjects to resolve the problem.
Both images are required to present comparable features on top of which
  an objective function is built to evaluate fitness of the solution.
However, oftentimes one of the images is not available or the objective
  function is not adequate or numerically tractable.

We propose a surface-to-volume variational solution to the registration problem,
  that exploits the detailed anatomy extracted manually or automatically
  from structural images.
We use an ``active contours without edges'' model to search for a
  deformation field of B-Spline functions that optimally maps the
  surface prior onto the data in the target space.
The strategy implicitly segments with sub-pixel resolution the image.
We illustrate the usefulness of the approach on a digital phantom.
Finally, we apply the registration to correct for suscetibility distortions
  10 real diffusion MRI datasets.
The joint segmentation-registration method successfully estimated the
  distortions.
We also compare its performance to a widely-used correction method that
  demands for an extra MR acquisition.