### 2015-02-27 Publications/2015-NeuroImage/supplemental.tex
paper version candidate

% @Author: Oscar Esteban
% @Date:   2015-02-20 15:40:09
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-02-27 15:21:03

\documentclass[a4paper]{report}

\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}

\usepackage{aliascnt}
\usepackage{relsize}
%\usepackage{inconsolata}
\usepackage[scaled]{beramono}
\usepackage[procnames]{listings}

\usepackage{natbib}
%\usepackage[style=numeric-comp,doi=false]{biblatex}

% *** GRAPHICS RELATED PACKAGES ***
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{ifpdf}
\graphicspath{{figures/},{../Figures/}}

% *** MATH PACKAGES ***
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}

\usepackage[T1]{fontenc}
\usepackage{charter}
%\usepackage[expert]{mathdesign}
%\usepackage[libertine,cmintegrals,cmbraces,vvarbb]{newtxmath}

% *** SPECIALIZED LIST PACKAGES ***
%\usepackage{algorithmic}
%\usepackage{algorithm}

% *** ALIGNMENT PACKAGES ***
%\usepackage{array}

% *** SUBFIGURE PACKAGES ***
\usepackage[font={small}]{caption}
%\usepackage{sidecap}

% *** FLOAT PACKAGES ***
\usepackage[framemethod=tikz]{mdframed}
%\usepackage{float}
% \usepackage{fixltx2e}
% \usepackage{stfloats}
% \usepackage{dblfloatfix}

% *** MISC UTILITY PACKAGES ***
%\usepackage{fancyhdr}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs} % to use \toprule
\usepackage{makeidx}
%\usepackage[scale=2.0]{ccicons}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[marginparwidth=1.2in]{geometry}
\usepackage{changepage}
\usepackage{epigraph}
\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
\usepackage[hyphens]{url}
\usepackage[breaklinks,hidelinks]{hyperref}
\usepackage[hyphenbreaks]{breakurl} %fixes boxes spanning through pages

\renewcommand\thesection{S\arabic{section}}
\renewcommand\thesubsection{S\arabic{section}.\arabic{subsection}}

% Listings style -----------------------------------------------------------------------------
\definecolor{pink}{RGB}{255,0,90}
\definecolor{comments}{RGB}{50,120,110}
\definecolor{string}{RGB}{160,0,0}
\definecolor{keywords}{RGB}{0,150,0}
\definecolor{listingbg}{gray}{0.95}

\lstset{basicstyle=\ttfamily\smaller\relax}

\newcommand*{\codeinline}[1]{\colorbox{listingbg}{\lstinline!#1!}}
%\newcommand\codeinline{\lstinline}

\lstnewenvironment{bashcode}[1][]
  {\lstset{language={}}\lstset{%
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  breaklines=true,
  basicstyle=\ttfamily\smaller\relax,
  keywordstyle=\color{keywords}\bfseries,
  commentstyle=\color{comments}\itshape,
  stringstyle=\color{string},
  showstringspaces=false,
  %identifierstyle=\color{green},
  morekeywords={regseg},
  numbers=none,
  frame=single,
  xleftmargin= 10pt,
  xrightmargin= 10pt,
  framexleftmargin=10pt,
  frameround=tttt,
  fillcolor=\color{listingbg},
  backgroundcolor=\color{listingbg}
}}
{}

\lstnewenvironment{pythoncode}[1][]
  {\lstset{language=bash}\lstset{%
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  breaklines=true,
  basicstyle=\ttfamily\smaller\relax,
  keywordstyle=\color{keywords}\bfseries,
  commentstyle=\color{comments}\itshape,
  stringstyle=\color{string},
  showstringspaces=false,
  identifierstyle=\color{black!70},
  procnamekeys={def,class},
  morekeywords={models, lambda, forms, as, from, import},
  numbers=left,
  numberstyle=\smaller\color{black!60},
  stepnumber=1,
  numbersep=5pt,
  frame=single,
  xleftmargin= 40pt,
  xrightmargin= 20pt,
  framexleftmargin=20pt,
  frameround=tttt,
  fillcolor=\color{gray!10},
  backgroundcolor=\color{gray!10}
}}
{}

\begin{document}
\title{Supplemental Materials: \emph{Simultaneous segmentation and registration of
diffusion MR images of the brain driven by active-contours}}
\author{Oscar Esteban}
\date{February 2015}

\maketitle

\section{Parameter settings and implementation details of \emph{regseg}}

\subsection{Implementation}
\paragraph{General} The \emph{regseg} registration and segmentation tool is written in C++, using ITK-4.6 as
  implementation core.
We designed a modular implementation, enabling multithreading in several pieces of the software,
  as the process is computationally expensive.
The tool generates a log-file in JSON format to easily inter-operate with secondary tools (such
  as the convergence report generation, \autoref{sec:convergence_evidence}).

\paragraph{Efficient interpolation using sparse matrices}
During the registration process, every iteration requires computing the product of all the speeds
  $\mathbf{s}'_i$ computed at the current position $\mathbf{v}'_i$ by the corresponding weights
  $\psi_k(\mathbf{v}_i)$ of interpolating functions (Eq. (15)).
As these weights can be computed once in the beginning of the process and they do not change along
  it, and they are different from zero only in the proximity of the corresponding control node $k$,
  they can be pre-cached in a sparse matrix.

\subsection{Interface and Settings}\label{sec:interface_settings}

\paragraph{Command-line interface}
The command line interface of \emph{regseg} supports general settings and level-wise settings.
For each multi-resolution level, its corresponding settings are added between brackets.

\begin{bashcode}
regseg -F fa.gz adc.nii.gz -P white.vtk pial.vtk -o myprefix [ -a 0.00000 -b 0.00000 --convergence-energy -t 1.0e-06 -w 60 --adaptative-descriptors --grid-spacing 16.0 -i 500 -s 0.001] [ -a 0. -b 0. --convergence-energy -t 1.e-08 -w 5 --grid-spacing 8.0 -i 250 -s 0.01]
\end{bashcode}

It is possible to get the description of available options running
  \codeinline{regseg -h}:

\begin{bashcode}
Usage:

General options:
  -h [ --help ]                         show help message
  -F [ --fixed-images ] arg             fixed image file
  -P [ --surface-priors ] arg           shape priors
  -T [ --surface-target ] arg           final shapes to evaluate metric (only
                                        testing purposes)
  -M [ --fixed-mask ] arg               fixed image mask
  -L [ --transform-levels ] arg         number of multi-resolution levels for
                                        the transform
  -o [ --output-prefix ] arg (=regseg)  prefix for output files
  -l [ --logfile ] arg                  log filename
  -v [ --monitoring-verbosity ] arg (=1)
                                        verbosity level of intermediate results
                                        monitoring ( 0 = no output; 5 = verbose
                                        )

Optimizer options (by levels):
  -a [ --alpha ] arg              alpha value in regularization
  -b [ --beta ] arg               beta value in regularization
  -s [ --step-size ] arg          step-size value in optimization
  -g [ --gradient-scales ] arg    alpha value in regularization
  -r [ --learning-rate ] arg      learning rate to update step size
  -i [ --iterations ] arg         number of iterations
  -w [ --convergence-window ] arg number of iterations of convergence window
  -t [ --convergence-thresh ] arg convergence value
  --grid-size arg                 size of control points grid
  --grid-spacing arg              spacing between control points
  -u [ --update-descriptors ] arg frequency (iterations) to update descriptors
                                  of regions (0=no update)
  --adaptative-descriptors        recomputes descriptors more often at the
                                  beginning of the process
  --convergence-energy            disables lazy convergence tracking: instead
                                  of fast computation of the mean norm of the
                                  displacement field, it computes the full
                                  energy functional

Functional options (by levels):
  --smoothing arg               apply isotropic smoothing filter on target
                                image, with kernel sigma=S mm.
  --smooth-auto                 apply isotropic smoothing filter on target
                                image, with automatic computation of kernel
                                sigma.
  --uniform-bg-membership       consider last ROI as background and do not
                                compute descriptors.
  -d [ --decile-threshold ] arg set (decile) threshold to consider a computed
                                gradient as outlier (ranges 0.0-0.5)

\end{bashcode}

\paragraph{\emph{Nipype} interface}
Our registration algorithm is released with a \emph{nipype Interface} packaged in
  \codeinline{pyacwereg.interfaces.acwereg}.
This interface has been comprehensively used in the evaluation workflows.

\begin{pythoncode}
from pyacwereg.interfaces.acwereg import ACWEReg
regseg = ACWEReg()
regseg.inputs.in_fixed = ['T1w.nii.gz', 'T2w.nii.gz']
regseg.inputs.in_pior = ['csf.vtk', 'white_lh.vtk', 'white_rh.vtk',
                         'pial_lh.vtk', 'pial_rh.vtk']
ifresult = regseg.run()
\end{pythoncode}

\subsection{Convergence evidencing}\label{sec:convergence_evidence}
In order to track the evolution of the registration process, several internal variables
  are saved in the JSON log-file.
Using the JSON log-file as input for the \emph{nipype Interface}
  \codeinline{ACWEReport}, it is straightforward to obtain
  a visual assessment document presenting the convergence.
Online checking is also possible as the algorithm writes to the standard output as well.

\begin{pythoncode}
from pyacwereg.interfaces.acwereg import ACWEReport
report = ACWEReport()
report.inputs.in_log = `myprefix.log'
ifresult = report.run()
\end{pythoncode}

\begin{figure*}[!ht]
	\includegraphics[width=\textwidth]{figures/Suppl-figure02.pdf}
	\caption{The evolution of the registration and segmentation process can be
	  checked using the \emph{Convergence report},
	  easily generated using the appropriate \emph{nipype Interface}.
	The report comprehends several plots tracking the evolution of the algorithm and several
	  features to help researchers tune up the algorithm in their application.}%
	  \label{fig:convreport}
\end{figure*}

\section{Instruments for evaluation}

This work is supported by two \emph{nipype Workflows} in order to ensure the reproducibility
  of the results.
All the intermediate results and figures in this paper have been encapsulated into
  the workflows and are available in \url{http://dx.doi.org/zenodolink}.

An overview of the general workflow is presented in the Figure 2 of the paper.
However, this figure has been extremely simplified for the best of visualization.
In this section, we review the main elements of the evaluation pipelines.

\paragraph{Preprocessing}

\paragraph{Distortion}

\section{Model considerations}

\begin{figure*}[!ht]
	\includegraphics[width=\textwidth]{figures/Suppl-figure01.pdf}
	\caption{Evaluating the joint distribution}\label{fig:jointplot}
\end{figure*}

\section{Extended results on phantom data}

\section{Extended results on real data}

\end{document}

%{\color{red} {The main diverging points with respect to
%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
%but we replace the Gaussian filtering by other constraints
%studied in \citep{nagel_investigation_1986} to better the problem;
%3) optimization is applied in the spectral
%domain, observing anisotropic and inhomogeneous mappings along each direction.
%With respect \citep{guyader_combined_2011}, the main differences are
%the distance function, and the spectral solution to the optimization updates,
%as we shall cover in \autoref{sec:numerical_implementation}.}}

%In this paper we formulate the joint registration-segmentation
%problem as follows.
%such that the known contours in anatomical space $T$ optimally segment
%the diffusion space $D$.
%
%
%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
%we alternatively use \emph{shape-gradients}
%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.

%
%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
%  its associated feature vector in the following.
%
%In our application, these surfaces are precise tissue interfaces of interest extracted
%  from a high-resolution, anatomically correct reference volume using
%  the well-established
%
%In this paper we propose a novel registration framework to simultaneously
%solving the segmentation, distortion and cortical parcellation challenges,
%by exploiting as strong shape-prior the detailed morphology extracted
%from high-resolution and anatomically correct \gls{mri}.
%Indeed, hereafter
%we assume this segmentation problem in anatomical images is reliably and
%accurately solved with readily available tools (e.g.
%\citep{fischl_freesurfer_2012}).
%After global alignment with \gls{t1} using existing approaches, the remaining
%spatial mismatch between anatomical and diffusion space is due to susceptibility
%distortions.
%Finally, we need to establish precise spatial correspondence between the
%surfaces in both spaces, including the tangential direction for parcellation.
%Therefore, we can reduce the problem to finding the differences of spatial
%distortion in between anatomical and \gls{dwi} space.
%We thus reformulate the segmentation problem as an inverse problem, where we
%seek for an underlying deformation field (the distortion) mapping
%from the structural space into the diffusion space, such that the structural
%contours segment optimally the \gls{dwi} data.
%In the process, the one-to-one
%correspondence between the contours in both spaces is guaranteed, and projection
%of parcellisation to \gls{dwi} space is implicit and consistent.
%
%We test our proposed joint segmentation-registration model on two different
%synthetic examples.
%The first example is a scalar sulcus model, where the
%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
%segmented correctly thanks to the shape prior and its coupling with the inner,
%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
%The second case deals with more realistic \gls{dwi} data stemming from
%phantom simulations of a simplistic brain data.
%Again, we show that the
%proposed model successfully segments the \gls{dwi} data based on two derived
%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
%of the dense distortion field.
%
%The rest of this paper is organized as follows.
%First, in \autoref{sec:methods}
%we introduce our proposed model for joint multivariate segmentation-registration.
%Then we provide a more detailed description of the data and experimental setup in
%\autoref{sec:experiments}.
%We present results in \autoref{sec:results} and conclude
%in \autoref{sec:conclusion}.
%
%
%The properties of the reconstructed tensors and derived scalar maps have
%been studied by \cite{ennis_orthogonal_2006}. Based on their
%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
%considered complementary features, and therefore we selected them for the
%energy model \eqref{eq:complete_energy} in driving the
%registration-segmentation process.
%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
%the \gls{md} is more related to the \emph{magnitude} of the process:
%
%\begin{align}
%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
%\end{align}
%where $\lambda_i$ are the eigenvalues of the diffusion tensor
%associated with the diffusion signal $S(\vec{q})$. There exist
%two main reasons to justify their choice.
%First, they are well-understood and standardized in clinical routine.
%Second, together they contain most of the information that is
%usually extracted from the \gls{dwi}-derived scalar maps
%\cite{ennis_orthogonal_2006}.
%