### 2015-04-06 Publications/2015-NeuroImage/00-main.tex
change kw

\documentclass[5p,authoryear]{elsarticle}
% \documentclass[authoryear,onecolumn]{elsarticle}
\journal{ArXiv archiving for pre-publication review}

% *** MISC UTILITY PACKAGES ***
\usepackage{glossaries}
\usepackage{booktabs} % to use \toprule
\usepackage{multirow}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xifthen}
\usepackage{xspace}

%\usepackage[cmex10,intlimits]{amsmath}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{algorithmic}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
\providecommand{\omegaset}{\ensuremath{\boldsymbol{\Omega}}}
\providecommand{\gammaset}{\ensuremath{\boldsymbol{\Gamma}}}

\let\oldhat\hat
\renewcommand{\vec}[1]{\mathbf{#1}}

\providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
\providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
\providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2%
\ifthenelse{\equal{#1}{}}{}{ [#1]}\xspace}

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{doi}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer %
align-ed pre-sent work-flow data-set data-sets}

\input{01-acronyms}

\usepackage[T1]{fontenc}
\usepackage{charter}
%\usepackage[expert]{mathdesign}
%\usepackage[libertine,cmintegrals,cmbraces,vvarbb]{newtxmath}
%\renewcommand*{\familydefault}{\sfdefault}
\usepackage[switch]{lineno}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% http://tex.stackexchange.com/questions/56105/missing-references-heading
\def\bibsection{\section*{References}}
\begin{document}
\begin{frontmatter}

%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
% \dochead{Preprint for ArXiv archiving and pre-submission review}

\title{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}

\author[bit,ciber]{Oscar~Esteban\corref{corrauthor}}
\cortext[corrauthor]{Corresponding author}
\ead{phd@oscaresteban.es}
\author[ucla]{Dominique~Zosso}
\author[scil,lts5]{Alessandro~Daducci}
\author[chuv,lts5]{Meritxell~Bach-Cuadra}
\author[bit,ciber]{Mar\'ia-J.~Ledesma-Carbayo}
\author[lts5]{Jean-Philippe~Thiran}
\author[bit,ciber]{Andres~Santos}

\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, %
Universidad Polit\'ecnica de Madrid, Madrid, Spain}
\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
\address[ucla]{Department of Mathematics, University of California,
Los Angeles (UCLA), Los Angeles, CA, US}
\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\address[chuv]{Dept. of Radiology, University
Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}

\begin{abstract}
\input{02-abstract}
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{keyword}
diffusion MRI \sep cortical~parcellation \sep segmentation %
\sep nonlinear~registration \sep active~contours \sep susceptibility~distortion.
\end{keyword}

\end{frontmatter}

\linenumbers
\input{03-introduction}
\input{04-methods}
\input{05-results}
\input{06-discussion}

\input{07-conclusion}

% ACKs
\input{08-acknowledgements}

% Add appendices
\appendix
\numberwithin{equation}{section}
\input{08-appendix-00-shapepriors}

%\bibliographystyle{plainnat}
\bibliographystyle{mystyle}
\bibliography{Remote}

\end{document}