### 2015-04-09 Publications/2015-NeuroImage/08-acknowledgements.tex
updates and standalone figures

% -*- root: 00-main.tex -*-
% use section* for acknowledgement
\section*{Author contributions}
All the authors contributed in editing the paper.
OE implemented the method, designed and conducted the experiments, wrote the paper,
  simulated the phantoms and prepared the real data.
DZ devised and drafted the registration method, generated early phantom datasets, and
  collaborated with the implementation of the method.
AD, MBC and MJLC interpreted the results.
AD, MBC, MJLC, JPT and AS advised on all aspects of the work.

\section*{Acknowledgments}
The authors thank Y. Alem\'an and G. Wollny for their thorough reviews of the manuscript,
  V. Estellers for early discussions at the beginning of this project,
  and L. A. Vese for her support during OE's research visits in her laboratory.
We also thank A. Leemans for kindly sharing a p-code version of \emph{ExploreDTI} from
  which settings for \emph{elastix} could be extracted.

DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727 
  and P300P2-147778, and NSF-DMS 1418812
This study is supported by: the Spain's Ministry of Science and Innovation
  (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
  European Regional Development Funds; the Center for Biomedical Imaging
  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
  Leenaards and Louis Jeantet foundations.