### 2015-04-09 Publications/2015-NeuroImage/figure05.tex

% -*- root: 00-main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-04-09 13:06:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-04-09 15:52:17

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
\begin{tikzpicture}
  \node[inner sep=10pt, draw=black!40](fig03b) at (0,8.8)
    {\hspace*{20pt}\includegraphics[width=21cm]{figure05A}};
  \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
    {\hspace*{20pt}\includegraphics[width=21cm]{figure05B}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,12.0) {\Huge \textbf{A}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,4.05) {\Huge \textbf{B}};
\end{tikzpicture}
\end{document}