### 2015-09-08 Publications/2015-NeuroImage/figure02-model.tex
v1.0 of resubmission

% -*- root: 00-main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-04-09 13:06:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-09-08 13:12:00

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
\begin{tikzpicture}
  \node[inner sep=10pt](fig01a) at (0,15.5)
    {\hspace*{20pt}\includegraphics[width=20cm]{figure04A}};
  \node[inner sep=10pt](fig01b) at (0,7.8)
    {\hspace*{20pt}\includegraphics[width=20cm]{figure04B}};
  \node[inner sep=10pt](fig01c) at (0,0)
    {\hspace*{20pt}\includegraphics[width=20cm]{figure04C}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,18.5) {\Huge \textbf{A}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,11.0) {\Huge \textbf{B}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,3.15) {\Huge \textbf{C}};
\end{tikzpicture}
\end{document}