### 2015-10-26 Publications/2015-NeuroImage/00-main.tex
v1.3.0rc

%\documentclass[5p,authoryear,fleqn]{elsarticle}
\documentclass[3p,authoryear,fleqn]{elsarticle}
\journal{NeuroImage}
\usepackage[utf8]{inputenc}

\usepackage{xifthen}
\newboolean{review}
\setboolean{review}{true}

% *** MISC UTILITY PACKAGES ***
\usepackage[mode=buildnew]{standalone}
\usepackage{gincltex}
\usepackage{glossaries}
\usepackage{booktabs} % to use \toprule
\usepackage{multirow}

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
\usepackage[colorlinks=true]{hyperref}
% \usepackage[hyperref]{xcolor}
\usepackage{doi}

\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{figures/}}

%\usepackage[cmex10,intlimits]{amsmath}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{algorithmic}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}
\providecommand{\e}[1]{\ensuremath{\cdot\text{10}^\text{#1}}}
\providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
\providecommand{\omegaset}{\ensuremath{\boldsymbol{\Omega}}}
\providecommand{\gammaset}{\ensuremath{\boldsymbol{\Gamma}}}
\providecommand{\regseg}{\emph{regseg}}
\providecommand{\Regseg}{\emph{Regseg}}
\providecommand{\lowb}{\textit{b0}}

\let\oldhat\hat
\renewcommand{\vec}[1]{\mathbf{#1}}

\providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
\providecommand{\suppl}[1]{\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{Supplemental Materials}, #1}
\providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2%
\ifthenelse{\equal{#1}{}}{}{ [#1]}\xspace}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer %
align-ed pre-sent work-flow data-set data-sets micro-struc-tu-re}

\input{01-acronyms}

\usepackage[T1]{fontenc}
\usepackage{charter}
%\usepackage[expert]{mathdesign}
%\usepackage[libertine,cmintegrals,cmbraces,vvarbb]{newtxmath}
%\renewcommand*{\familydefault}{\sfdefault}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{ifpdf}

\ifpdf
  \usepackage[switch]{lineno}
  % nothing to do
\else
  \makeatletter
  \def\temp{dvips.def}
  \ifx\Gin@driver\temp
  \def\Ginclude@graphics#1{\def\temp{#1}---image \expandafter\strip@prefix\meaning\temp---}
  \fi
  \makeatother
\fi

\ifthenelse{\boolean{review}}
{
  \usepackage[colorinlistoftodos,textsize=tiny, textwidth=1.2cm]{todonotes}
  \setlength{\marginparwidth}{1.5cm}
  \reversemarginpar
  % \setlength{\marginparwidth}{2.0cm}
  \newcommand{\revcomment}[2][]{\ifthenelse{\equal{#1}{}}{}{\todo{#1}}%
  \begingroup \color{blue!70} #2\endgroup}
}
{
  \newcommand{\todo}[2][]{}
  \newcommand{\revcomment}[2][]{#2}
}

% http://tex.stackexchange.com/questions/56105/missing-references-heading
\def\bibsection{\section*{References}}

\begin{document}
\hypersetup{linkcolor=black!60, citecolor=black!60, urlcolor=black!60}
\begin{frontmatter}

%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
% \dochead{Preprint for ArXiv archiving and pre-submission review}

\title{Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}}
\tnotetext[t1]{Supplemental materials to this work will be publicly available upon publication with doi:\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{10.6084/m9.figshare.1397502}.
During the review process, the document is available with the private link.}

\author[bit,ciber]{Oscar~Esteban\corref{corrauthor}}
\cortext[corrauthor]{Corresponding author}
\ead{phd@oscaresteban.es}
\author[ucla]{Dominique~Zosso}
\author[lts5]{Alessandro~Daducci}
\author[chuv,lts5]{Meritxell~Bach-Cuadra}
\author[bit,ciber]{Mar\'ia-J.~Ledesma-Carbayo}
\author[lts5]{Jean-Philippe~Thiran}
\author[bit,ciber]{Andres~Santos}

\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain}
\address[ucla]{Department of Mathematics, University of California,
Los Angeles (UCLA), Los Angeles, CA, US}
% \address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\address[chuv]{Dept. of Radiology, CIBM, University
Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}

\begin{abstract}
\input{02-abstract}
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{keyword}
active~surfaces \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
\end{keyword}

\end{frontmatter}

\linenumbers
\input{03-introduction}
\input{04-methods}
\input{05-results}
\input{06-discussion}

\input{07-conclusion}

% ACKs
\input{08-acknowledgements}

% Add appendices
\appendix
\numberwithin{equation}{section}
\input{08-appendix-00-shapepriors}

%\bibliographystyle{plainnat}
\bibliographystyle{mystyle}
\bibliography{Remote}

\end{document}