### 2015-10-29 Publications/2015-NeuroImage/03-introduction.tex

% -*- root: 00-main.tex -*-
\section{Introduction}\label{sec:regseg-intro}
\Acrlong*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
  and connectivity \citep{craddock_imaging_2013} of the human brain \emph{in-vivo}.
It is generally acquired using \gls*{epi} schemes, since they are very fast at
  scanning a large sequence of images called \glspl*{dwi}.
Each \gls*{dwi} is sensitized with a gradient to probe proton diffusion in a certain
  \revcomment[C13 (R.2)]{orientation}.
Subsequent processing involves describing the local microstructure with one available
  model, which ranges from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
  to current models such as \citep{daducci_accelerated_2015}.
The microstructural map is then used to draw the preferential \revcomment[C13 (R.2)]{orientations} of diffusion
  across the brain using tractography \citep{mori_threedimensional_1999}.
Finally, a graph representing the corresponding structural network is built using
  the regions of a cortical parcellation as nodes and the fiber paths found by
  tractography as edges \citep{hagmann_mapping_2008}.
\revcomment[C9 (R.2) \\ C17 (R.2)]{%
The methodologies to solve reconstruction, tractography and network building
  require the delineation of the anatomy in the \gls*{dmri} space.
Moreover, current trends on reconstruction \citep{jeurissen_multitissue_2014} and
  tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
  structural information to improve the microstructural mapping and fiber-tracking.}

\revcomment[C15 (R.2)]{%
Possibly, the earliest structural information incorporated to aid \gls*{dmri} processing
  is the \gls*{wm} mask used as a termination criteria for tractography.
The standardized procedure to obtain this mask was thresholding the \gls*{fa} map.
However, the mask and subsequent analyses are highly dependent on the threshold set
  \citep{taoka_fractional_2009}.
\todo{C16 (R.2)}
To overcome the unreliability of \gls*{fa} thresholding, and to broaden
  \gls*{wm} segmentation to brain tissue segmentation, a large number of
  methods have been proposed using \glspl*{dwi}, the \lowb{}, and \gls*{dti}-derived
  scalar maps such as \gls*{fa}, \gls*{adc} and others \citep{zhukov_level_2003,
  rousson_level_2004,jonasson_segmentation_2005,hadjiprocopis_unbiased_2005,liu_brain_2007,
  lu_segmentation_2008,han_experimental_2009}.}
\todo{C14 (R.2)}%
However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
First, \gls{dmri} images have a resolution that is \revcomment[C12 (R.2)]{much lower}
  than that of the imaged microstructural features.
Therefore, voxels located in structural discontinuities are affected by partial
  voluming of the signal sources.
Second, the extremely low \gls*{snr} and the high dimensionality of the \glspl*{dwi} prevent
  their direct use in segmentation.
Third, the low contrast between \gls*{gm} and \gls*{wm} in the \lowb{} volume also makes
  it unsuitable for brain tissue segmentation.

\revcomment[C18 (R.2)]{%
An alternative route to segmentation in \gls*{dmri} space is the mapping of the
  structural information extracted from anatomical MR images, such as \gls*{t1}, using
  image registration techniques.
Generally, intra-subject registration of MR images of the brain involves only a linear
  mapping to compensate for head motion between scans.
However, \gls*{epi} introduces a geometrical distortion \citep{jezzard_correction_1995}
  that impedes the linear mapping from the structural space.
Numerous methods have been proposed to overcome this problem by incorporating information
  from extra MR acquisitions such as fieldmaps \citep{jezzard_correction_1995},
  \glspl*{dwi} with a different \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000},
  or \gls*{t2} images \citep{kybic_unwarping_2000,studholme_accurate_2000}.
These methods estimate the deformation field associated to \emph{\gls*{epi} distortions} and
  resample the \glspl*{dwi} onto a corrected \gls*{dmri} space.
The retrospective \gls*{epi} correction is an active field of research yielding frequent refinements
  and combinations of the original methods, such as
  \citep{holland_efficient_2010,andersson_comprehensive_2012,
  irfanoglu_drbuddi_2015}.
A standardized method to solve the remaining linear mapping between the corrected-\gls*{dmri}
  and the structural spaces is \emph{bbregister} \citep{greve_accurate_2009}.%
}

\revcomment[rephrased]{%
Here we present a segmentation and surface-to-volume registration method
  named \regseg{}, and show its usefulness in mapping anatomical information from structural
  into native \gls*{dmri} space to aid subsequent processing steps
  (reconstruction, tractography and network building using a cortical parcellation).
The underlying hypothesis is that the registration and segmentation problems
  in \gls*{dmri} can be solved simultaneously.}
\revcomment[C1 (R.1) \\ C10 (R.2)]{%
To implement \regseg{} we first establish an active-contours without edges
  \citep{chan_active_2001} segmentation framework.}
\revcomment[C2 (R.1)]{%
A specific set of reference surfaces extracted from the same subject initialize
  the 3D active-contours, which evolve searching for homogeneous regions in the multivariate
  target-image.
In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}
  using a set of nested surfaces extracted from a strutrucal image (e.g. \gls*{t1}) and
  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps as inputs.
The evolution of the surfaces is supported by a B-spline basis, optimized
  iteratively using a descent approach driven by shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, \regseg{} establishes a registration framework that actually
  overcomes the nonlinear warping induced by \gls*{epi} distortions.%
}
\Regseg{} integrates the benefits of segmentation and registration methods together and
  exploits the multivariate nature of \gls*{dmri} data to contribute in the proposed
  application on three key aspects:
  1) it avoids the resampling the \glspl*{dwi} introduced by the correction of \gls*{epi}
    distortions, since processing can be performed on the native \gls*{dmri} space;
  2) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
    therefore \regseg{} does not require additional MR acquisitions to the minimal
    \gls*{dmri} protocol; and
  3) \regseg{} increases the geometrical accuracy of the overall process.
In this paper, we first verify the functionality of the method and the \regseg{}
  implementation using a set of digital phantoms, demonstrating the sub-pixel accuracy
  in registration.
We then evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
  instrumentation framework \citep{esteban_simulationbased_2014} which simulates
  known and realistic \gls*{epi} distortions.
We also compare \regseg{} and a nonlinear registration method to map the \lowb{} to
  the corresponding \gls*{t2} image of the same subject.
This method is a derivation of the \gls*{t2b} correction methods \citep{kybic_unwarping_2000},
  reproducing the settings of an existing implementation included within a widely used diffusion
  processing software (\emph{ExploreDTI}, \cite{leemans_exploredti_2009}).
With this comparison, we demonstrate how \regseg{} achieves higher accuracy with the
  simultaneous registration and segmentation process.