### 2016-01-16 Publications/2015-NeuroImage/cover_letter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font
\usepackage{graphicx}
\usepackage[all]{background}

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

%\usepackage{showframe}
\usepackage{tikz}

\newcommand{\MyGraphicLogo}{% For imported graphic logo
\begin{tikzpicture}[remember picture,overlay,yshift=-2cm, xshift=4cm]
  \node at (0,0) {\includegraphics[height=1.80in]{/home/oesteban/Pictures/logocei.png}};
 \end{tikzpicture}
}

\SetBgContents{\MyGraphicLogo}% Select included image
\SetBgPosition{current page.north west}% Select location
\SetBgOpacity{1.0}% Select opacity
\SetBgAngle{0.0}% Select roation of logo
\SetBgScale{1.0}% Select scale factor of logo

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{\textbf{Stephen M. Smith}, \\
Senior Editor, \\
NeuroImage
}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\includegraphics[height=1.83in]{../Figures/logocei}

\begin{flushright}
\textbf{Oscar Esteban}, \\ % Your name
Researcher at Biomedical Image Technologies (BIT), \\
Universidad Polit\'ecnica de Madrid \\
Av. Complutense, 30 - ETSI Telecomunicaci\'on C203 \\
E-28040 Madrid (Spain) \\
Tlf.: +34 915 495 700 ext.4234,\\
E-mail: oesteban@die.upm.es % Your address and phone number
\end{flushright}
\vfill

\signature{Oscar Esteban} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Editor,}

Please, find enclosed our manuscript, \emph{Surface-driven registration method for the structure-informed segmentation
  of diffusion MR images}, by O. Esteban et al., which we would like to submit for publication in \emph{NeuroImage}.
\textbf{This work is a resubmission of the previously rejected paper, NIMG-15-1151}, as suggested in its
  corresponding decision letter.

For convenience, a point-by-point response for the reviewers of the submission NIMG-15-1151 is attached, and the
  paper is sent in two versions, one annotated for the reviewers and another without annotations.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.
The pre-print of the work has been made available in bioRxiv (doi 10.1101/018945), after checking there
  is no collision with the policies of NeuroImage.

We look forward to hearing from you at your earliest convenience.

\closing{Sincerely yours,}

%\encl{Curriculum vitae, employment form} % List your enclosed documents here, comment this out to get rid of the "encl:"

%----------------------------------------------------------------------------------------
\vfill
\end{letter}

\end{document}