### 2016-02-03 Publications/2015-NeuroImage/cover_letter2.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{charter}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font
\usepackage{graphicx}
\usepackage[all]{background}

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

%\usepackage{showframe}
\usepackage{tikz}

\newcommand{\MyGraphicLogo}{% For imported graphic logo
\begin{tikzpicture}[remember picture,overlay,yshift=-2cm, xshift=4cm]
  \node at (0,0) {\includegraphics[height=1.80in]{/Users/oesteban/Pictures/logocei.png}};
 \end{tikzpicture}
}

\SetBgContents{\MyGraphicLogo}% Select included image
\SetBgPosition{current page.north west}% Select location
\SetBgOpacity{1.0}% Select opacity
\SetBgAngle{0.0}% Select roation of logo
\SetBgScale{1.0}% Select scale factor of logo

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{\textbf{Stephen M. Smith}, \\
Senior Editor, \\
NeuroImage
}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\includegraphics[height=1.83in]{../Figures/logocei}

\begin{flushright}
\textbf{Oscar Esteban}, \\ % Your name
Researcher at Biomedical Image Technologies (BIT), \\
Universidad Polit\'ecnica de Madrid \\
Av. Complutense, 30 - ETSI Telecomunicaci\'on C203 \\
E-28040 Madrid (Spain) \\
Tlf.: +1 (707) 492 61 40,\\
E-mail: \texttt{phd@oscaresteban.es} % Your address and phone number
\end{flushright}
\vfill

\signature{Oscar Esteban} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Editor,}

Please, find enclosed our manuscript, \emph{Surface-driven registration method for the structure-informed segmentation
  of diffusion MR images}, by O. Esteban et al., which we would like to submit for publication in \emph{NeuroImage}.
This manuscript \textbf{is a revision of the submission NIMG-15-2877}.
For convenience, a point-by-point response for the reviewers is attached.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.
The pre-print of the work has been made available in bioRxiv (doi 10.1101/018945), after checking there
  is no collision with the policies of NeuroImage.

We look forward to hearing from you at your earliest convenience.

\closing{Sincerely yours,}

%\encl{Curriculum vitae, employment form} % List your enclosed documents here, comment this out to get rid of the "encl:"

%----------------------------------------------------------------------------------------
\vfill
\end{letter}

\end{document}