### 2016-03-10 Publications/2015-NeuroImage/figure06.tex

% -*- root: 00-main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-04-09 14:39:30
% @Last Modified by:   oesteban
% @Last Modified time: 2016-03-09 13:48:53

\documentclass[tikz]{standalone}

\thispagestyle{empty}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
  \begin{tikzpicture}
    \node[inner sep=15pt, draw=black!40](fig06a) at (0,12.0)
      {\hspace*{20pt}\includegraphics[width=21cm]{figure06A}};
    \node[inner sep=15pt, draw=black!40](fig06b) at (0,0)
      {\hspace*{20pt}\includegraphics[width=21cm]{figure06B}};
    \node[circle, text=black!75, inner sep=0] at (-10.65,18.6) {\Huge \textbf{A}};
    \node[circle, text=black!75, inner sep=0] at (-10.65,3.5) {\Huge \textbf{B}};
  \end{tikzpicture}
\end{document}