### 2016-03-10 Publications/2015-NeuroImage/figure02.tex

% @Author: Oscar Esteban
% @Date:   2015-04-09 13:06:32
% @Last Modified by:   oesteban
% @Last Modified time: 2016-03-10 09:23:55

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
\noindent
\begin{minipage}[t]{14in}
\noindent
\begin{tikzpicture}[block/.style={}]
  \node[inner sep=0pt](fig02a) at (0,14)
    {%
    \noindent
    \begin{minipage}[t]{.2\textwidth}
    \centering
	\vskip 20pt
    \includegraphics[width=\textwidth]{figure02A-1}
    \vskip 10pt
    \parbox{.9\textwidth}{%
    \textbf{Reference}: segmentation by the surfaces
    extracted in T1w and original (undistorted)
    diffusion data.}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{.2\textwidth}
	\vskip 20pt
    \centering
    \includegraphics[width=\textwidth]{figure02A-2}
    \vskip 10pt
    \parbox{.9\textwidth}{%
    \textbf{Before registration}, the
	contours from the T1w are
	not aligned with the
	distorted diffusion data.}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{.2\textwidth}
	\vskip 20pt
    \centering
    \includegraphics[width=\textwidth]{figure02A-3}
    \vskip 10pt
    \parbox{.9\textwidth}{%
    \textbf{After registration}, the
	contours are now aligned
	with the distorted diffusion
	data.}
    \end{minipage}};
  \node[text=black!75, inner sep=0] at (-4.7,18.7) {\Large \textbf{A. OPTIMIZATION OF THE SEGMENTATION MODEL}};
  % \node[inner sep=10pt, draw=black!40](fig05c) at (8.,0)
  %   {\hspace*{20pt}};
  % \node[circle, text=black!75, inner sep=0] at (-10.65,4.05) {\Huge \textbf{B}};
  %
  %
  %
  %
  \node[inner sep=0pt](fig02b) at (0,0)
    {%
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-101}
    	\includegraphics[width=\textwidth]{figure02B-102}
    	\includegraphics[width=\textwidth]{figure02B-103}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-201}
    	\includegraphics[width=\textwidth]{figure02B-202}
    	\includegraphics[width=\textwidth]{figure02B-203}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-301}
    	\includegraphics[width=\textwidth]{figure02B-302}
    	\includegraphics[width=\textwidth]{figure02B-303}
    	\end{minipage}
	    \hskip 15pt
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-104}
    	\includegraphics[width=\textwidth]{figure02B-105}
    	\includegraphics[width=\textwidth]{figure02B-100}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-204}
    	\includegraphics[width=\textwidth]{figure02B-205}
    	\includegraphics[width=\textwidth]{figure02B-200}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-304}
    	\includegraphics[width=\textwidth]{figure02B-305}
		\includegraphics[width=\textwidth]{figure02B-300}
    	\end{minipage}%
	};
  \node[inner sep=3pt, draw=black!20](fig02c) at (0,-6.4)
    {\includegraphics[width=.57\textwidth]{legend}};
  \node[text=black!75, inner sep=0] at (-4.7,7.0) {\Large \textbf{B. DETAIL OF TISSUE-WISE JOINT-DISTRIBUTIONS}};
  \node[text=black!75, inner sep=0] at (-9.0,6.3) {\textbf{Reference}};
  \node[text=black!75, inner sep=0] at (-5.5,6.3) {\textbf{Pre-registration}};
  \node[text=black!75, inner sep=0] at (-2.0,6.3) {\textbf{Post-registration}};
  \node[text=black!75, inner sep=0] at (1.8,6.3) {\textbf{Reference}};
  \node[text=black!75, inner sep=0] at (5.4,6.3) {\textbf{Pre-registration}};
  \node[text=black!75, inner sep=0] at (8.8,6.3) {\textbf{Post-registration}};

  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (newmode0) at (4.8, 4.0) {New mode};
  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (newmode1) at (-5.6, -4.0) {New mode};
  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (spread0) at (4.0, -5.1) {Larger spread};
  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (spread1) at (-5.6, 0.0) {Larger spread};
  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (spread2) at (-5.6, 4.0) {Larger spread};

    \draw[-stealth, line width=2pt, black]
    (newmode0.south) -- ++(-0.85, -0.5);
    \draw[-stealth, line width=2pt, black]
    (newmode1.west) -- ++(-1.1, 0.6);
    \draw[-stealth, line width=2pt, black]
    (spread0.north east) -- ++(.2, .2);
    \draw[-stealth, line width=2pt, black]
    (spread1.south) -- ++(-0.7, -0.8);
    \draw[-stealth, line width=2pt, black]
    (spread2.south) -- ++(0, -1.25);
    \draw[-stealth, line width=2pt, black]
    (spread2.west) -- ++(-1.25, -0.1);
\end{tikzpicture}
\end{minipage}
\end{document}